---
title: "Preprocessing"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
# Loading library here
library(tidyverse)
library(biomformat)
library(qiime2R)
library(knitr)


# set path
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source("~/Dropbox/42-JCHWANG-RUTGERS/Projects-Rutgers/Src/utils.R")
setwd(rprojroot::find_rstudio_root_file())

path_ana <- "~/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing"

sessionInfo()
```


# 2020-04-15 Fix issues with mapping file
In the original mapping file, some subject specific information was inconsistent across different body sites from the same subject.

Therefore, both manual and automatical fix has been performed.

```{r}
Mapping_VL <- read.delim("../Mapping_Bolivar_Villagers-20181108.txt",stringsAsFactors = F)
# (md5): `47c35f47f6e9ce45e0bf9c052525eee2`


Mapping_VL[is.na(Mapping_VL)] <- "NA"

## remove village "Kadanskaduinha" becaue very few subjects, and those without age information
Mapping_MS_VL <- Mapping_VL %>% filter(Village != "Kadansakaduinha", Age!="NA")
Mapping_MS_VL$Ethnicity[which(Mapping_MS_VL$Ethnicity=="Yekwana")] <- "YEKWANA"
```

Identify which subjects have problems
```{r}
Mapping_VL_subject <- Mapping_MS_VL %>% select(-c(1:6, 9:10, 13, 14, 64:69, 88)) %>% distinct() %>% mutate(sub_unique = paste(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID, sep = "-"))

# check more than 1 records per subject
## Ethnicity, Village, House_ID, Family_ID, Subject_ID should identify an individual
Temp_more_records <- Mapping_VL_subject %>% group_by(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID) %>% filter(n()>1)
Temp_more_records_check <- Temp_more_records %>% group_by(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID, sub_unique) %>% summarise_all(n_distinct) %>% ungroup() %>% select_if(.predicate = function(x){any(x>1)})

Temp_tbl <- data.frame(Sub = character(), Col = character(), Value = character(), stringsAsFactors = F)
for (f in colnames(Temp_more_records_check)[8:61]){
  Temp <- Temp_more_records_check %>% filter(get(f)>1)
  Temp2 <- merge(Temp[, 1:7], Mapping_MS_VL, by = c("Year", "Ethnicity", "Village", "House_ID", "Family_ID", "Subject_ID"))
  Temp3 <- Temp2 %>% select(sub_unique, f) %>% group_by_at(vars(1, 2)) %>% summarise(N = n()) %>% mutate(Pct = N/sum(N), Value = paste(get(f), Pct, sep = ":")) %>% summarise(Value = paste(Value, collapse = ";")) %>% mutate(Col = f) %>% rename(Sub = sub_unique)
  Temp_tbl <- rbind(Temp_tbl, Temp3)
  # Temp2 <- merge(Mapping_MS_VL, Temp[, 1:6], by = c("Year", "Ethnicity", "Village", "House_ID", "Family_ID", "Subject_ID"))
  # browser()
}

```

Perform automatic revising of the mapping file after manual fixing some of the issues.
```{r eval=FALSE}
Work <- Mapping_MS_VL %>% mutate(sub_unique = paste(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID, sep = "-"))

# first fix the situation that there are only 2 values and 1 of them is NA
Work_d1 <- Work %>% group_by(sub_unique) %>% mutate_at(vars(c(46:87)), ~revise_para_1(.))
## check if it works
Work_d1_check_1 <- Work_d1 %>% select(-c(1:6, 9:10, 13, 14, 64:69, 88)) %>% distinct()
Work_d1_check_2 <- Work_d1_check_1 %>% group_by(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID) %>% filter(n()>1)
Work_d1_check_3 <- Work_d1_check_2 %>% group_by(Year, Ethnicity, Village, House_ID, Family_ID, Subject_ID, sub_unique) %>% summarise_all(n_distinct) %>% ungroup() %>% select_if(.predicate = function(x){any(x>1)})
## Yes, it works.

# Update Mapping_MS_VL
Mapping_MS_VL <- Work_d1 %>% ungroup() %>% select(-sub_unique)

# Update the original mapping
Mapping_VL <- rbind(Mapping_MS_VL, Mapping_VL %>% filter(!SampleID %in% Mapping_MS_VL$SampleID))
write.table(Mapping_VL, file = "../Mapping_Bolivar_Villagers-20200416-jw.txt", quote = F, row.names = F, sep = "\t")

revise_para_1 <- function(x){
    # this function is to revise the mapping file so that the same subject would have the same parasite result consistently, this function could only handle the situation when there are only 2 values and 1 of them is NA
    # argument:
    #   x, the vector of values for a specific column
    if (n_distinct(x)==2 & ("NA" %in% x)) {
        x[x=="NA"] <- x[x!="NA"][1]
    }
    return(x)
}

```

# Import mapping files
Mapping files that were used in this analysis are:

* `Mapping_Bolivar_Villagers-20200416-jw.txt` (md5): `209e8fd3fc0ffefc133b44d66c4f58ae`
* `Mapping_Bolivar_Visitor_2015.txt` (md5): `6c9acd9c8af33ba4b400af436c8d90aa`
* `Mapping_Bolivar_Visitor_2016.txt` (md5): `4a4ece643bd971989da75abcdf4dd0e2`

```{r}
Mapping_files <- paste0("../", c("Mapping_Bolivar_Villagers-20200416-jw.txt", "Mapping_Bolivar_Visitor_2015.txt", "Mapping_Bolivar_Visitor_2016.txt"))
Mapping_import <- lapply(Mapping_files, FUN = function(x){read.delim(x, stringsAsFactors = F)})
names(Mapping_import) <- c("VL", "VR15", "VR16")
```

## Create study specific mapping file
### Mapping_MS_Bolivar_Amerin1dians
(1) Included all Amerindians except for those who did not has Age information
(2) Included only the visitor's baseline samples
```{r}
Work <- Mapping_import
Work_VR15_BL <- Work$VR15 %>% group_by(Body_Site, Subject_ID) %>% filter(Days_of_study==min(Days_of_study)) %>% ungroup()
Work_VR16_BL <- Work$VR16 %>% group_by(Body_Site, Subject_ID) %>% filter(Days_of_study==min(Days_of_study)) %>% ungroup()

# House keeping
Work_VR15_BL$Ethnicity <- "Visitors"
Work_VR16_BL$Ethnicity <- "Visitors"
Work_VR15_BL$Village <- "Visitors"
Work_VR16_BL$Village <- "Visitors"
Work_VR15_BL$House_ID <- "Visitors"
Work_VR16_BL$House_ID <- "Visitors"
Work_VR15_BL$Family_ID <- "Visitors"
Work_VR16_BL$Family_ID <- "Visitors"


Mapping_common_header <- Reduce(intersect, list(colnames(Mapping_import$VL), colnames(Work_VR15_BL), colnames(Work_VR16_BL)))
Mapping_MS_Urban <- do.call("rbind", lapply(list(Mapping_import$VL, Work_VR15_BL, Work_VR16_BL), FUN = function(x){x[Mapping_common_header]})) %>% filter(Village != "Kadansakaduinha", !is.na(Age))
Mapping_MS_Urban$Ethnicity[which(Mapping_MS_Urban$Ethnicity=="Yekwana")] <- "YEKWANA"
save(Mapping_MS_Urban, file = "output/Mapping_MS_Urban.Rdata")
```


# Alpha diversity
```{r, eval=FALSE}
# These procedure has already been completed previously.
## QIIME2 alpha output
path1 <- "../4-Div_dada/export-alpha/"
path2 <- sapply(list.files(path1), FUN = function(x){paste0(path1, x, "/data/alpha-diversity.tsv")})
alpha_import <- lapply(path2, FUN = function(x){read.table(x, header = T, sep = "\t", row.names = 1, stringsAsFactors = F)})
alpha <- merge_iter(alpha_import, by = 0)
save(alpha, file = "output/alpha_imported_all-20190722.Rdata")
```

# Beta diversity

```{r, eval=FALSE}
path1 <- "../4-Div_dada/export-beta/"
path2 <- sapply(list.files(path1), FUN = function(x){paste0(path1, x, "/data/distance-matrix.tsv")})
beta_import <- lapply(path2, FUN = function(x){read.table(x, header = T, sep = "\t", row.names = 1, stringsAsFactors = F)})

# Unweighted unifrac
DM_uu <- beta_import[["unweighted_unifrac"]] %>% as.matrix() %>% as.dist()
save(DM_uu, file = "output/DM_uu_all_20180530.Rdata")

# Weighted unifrac
DM_wu <- beta_import$weighted_unifrac %>% as.matrix() %>% as.dist()
save(DM_wu, file = "output/DM_wu_all_20180530.Rdata")

DM_bc <- beta_import$bray_curtis %>% as.matrix() %>% as.dist()
save(DM_bc, file = "output/DM_bc_all_20200604.Rdata")
```

# Decoide
## run decoide
```{r}
load("output/Mapping_MS_Urban.Rdata")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")
for (BB in BodySites) {
    # BB = "Feces"
    for (YY in Yrs){
        # YY = "2015"
        Work_dat <- Mapping_MS_Urban %>% filter(Body_Site==BB, Year==YY)
        write.table(Work_dat, file = paste0("output/Mapping_MS_Urban_", BB, "_", YY, ".txt"), sep = "\t", quote = F, row.names = F)
    }
}
```

```{bash}
# qiime2-2019.7 with deicode installed
for f in /Users/Jincheng/Projects/bolivar_main_2020/output/Mapping_MS_Urban*.txt; do name=$(basename $f ".txt"); qiime feature-table filter-samples --i-table /Users/Jincheng/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/3-Dada/merged-otu.qza --m-metadata-file $f --o-filtered-table /Users/Jincheng/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/4.1-Deicode/${name}_ft.qza; done &>Deicode.log &

for f in /Users/Jincheng/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/4.1-Deicode/*_ft.qza; do name=$(basename $f "_ft.qza"); qiime deicode rpca --i-table $f --p-min-feature-count 10 --p-min-sample-count 500 --p-n-components 3 --o-biplot /Users/Jincheng/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/4.1-Deicode/${name}_ord.qza --o-distance-matrix /Users/Jincheng/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/4.1-Deicode/${name}_dist.qza; done &>>Deicode.log &
```

```{r, eval=FALSE}
path1 <- paste0(path_ana, "/4.1-Deicode/")
ord_qzas <- lapply(list.files(path = path1, pattern = "Mapping_MS_Urban.*ord.qza", full.names = T), read_qza)

names(ord_qzas) <- list.files(path = path1, pattern = "Mapping_MS_Urban.*ord.qza") %>% gsub(pattern = "Mapping_MS_Urban_|_ord.qza", replacement = "", .)

dist_qzas <- lapply(list.files(path = path1, pattern = "Mapping_MS_Urban.*dist.qza", full.names = T), read_qza)
names(dist_qzas) <- list.files(path = path1, pattern = "Mapping_MS_Urban.*dist.qza") %>% gsub(pattern = "Mapping_MS_Urban_|_dist.qza", replacement = "", .)

save(ord_qzas, file = "output/Ord_deicode_all_20200706.Rdata")
save(dist_qzas, file = "output/DM_deicode_all_20200604.Rdata")

```

# Feature table
```{r, eval = FALSE}
# Biom table
dat <- read_hdf5_biom("../4-Div_data/export-rarefied_table/data/feature-table.biom") # bddf0151-4eac-4835-9213-2867f2c70c74
biom <- biom(dat) %>% biom_data() %>% as.matrix() %>% as.data.frame()
save(biom, file = "output/FT_rarefied_full_11142019.Rdata")
```

