---
title: "Beta Diversity"
date: "`r Sys.Date()`"
output:
  html_document:
      theme: simplex
      code_folding: hide
      toc: true
      toc_depth: 2
      toc_float: true
      df_print: kable
editor_options:
  chunk_output_type: console
---

## Environ Setup
```{r setup}
# Loading library here
library(knitr)
library(tidyverse)
library(broom)
library(patchwork)
library(rstatix)
library(ape)
library(vegan)

options(max.print="75")
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, prompt=FALSE, tidy = TRUE, comment = NA, message = FALSE)
opts_knit$set(width=75)

source("assets/utils.R")

# number of cores/thread to use
nc <- switch(Sys.info()[['sysname']], Windows = 1, Darwin = 4, Linux = 12)

set.seed(20)
sessionInfo()
```

## Data Setup and Miscs
```{r}
#------- Setup color index ---------#
# Exposure level
Expo_color <- c("#1b9e77", "#d95f02", "#8da0cb")
names(Expo_color) <- c("Low", "Medium", "High")
print(Expo_color)
#plot(1:3, 1:3, col = Expo_color, pch = 16, cex = 4)
# Ethnicity
Ethnicity_color <- gg_color_hue(3)
names(Ethnicity_color) <- c("SANEMA", "YEKWANA", "Visitors")
print(Ethnicity_color)

Sanemas <- c("Chajuranha", "Mosenahanha", "Kuyuwininha", "Shianana-Jiyakwanha", "Washudihanha", "Sudukuma")
Yekwana <- c("Kanarakuni", "Fiyakwanha")

#------- Import mapping file---------## 
## This mapping files include all villagers and baseline of visitors of all body sites
mt_ms = readRDS("../data/Mapping_MS_Expo.Rds")
Mapping_work <- mt_ms %>% mutate(Expo = ifelse(Ethnicity=="SANEMA", "Low", ifelse(Village=="Fiyakwanha", "Low", ifelse(SampleGroup=="Visitors", "High", "Medium"))) %>% factor(levels = c("Low", "Medium", "High")), Age_num = as.numeric(Age), Age_num = ifelse(is.na(Age_num), as.numeric(gsub(pattern = "_months", replacement = "", Age, ignore.case = T))/12, Age_num), Age_grp1 = ifelse(Age_num>=18, "Adults", "Children"))
Mapping_work$Age_num[Mapping_work$Age=="4_Days"] <- 0
Mapping_work$Age_num[Mapping_work$Age=="1_day"] <- 0
Mapping_work$Age_grp1[Mapping_work$Age_num==0] <- "Children"

Mapping_work$Age_grp2 <- case_when(Mapping_work$Age_num<=3 ~ "Age_0-3", Mapping_work$Age_num<=8 ~ "Age_3-8", Mapping_work$Age_num<=18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults"))

#------- Import data ---------## 
# Beta distance matrix
## Unweighted unifrac
DM_uu <- readRDS("../data/DM_uu_all_20180530.Rds")
## Weighted unifrac
DM_wu <- readRDS("../data/DM_wu_all_20180530.Rds")
## bray curtis
DM_bs <- readRDS("../data/DM_bc_all_20200604.Rds")
## deicode
DM_dc <- readRDS("../data/DM_deicode_all_20200604.Rds")
Ord_dc <- readRDS("../data/Ord_deicode_all_20200706.Rds")

```

Due to analysis in alpha diversity, age 8 may be a good seperation piont, and also previous study seems to suggest age 3 also is a good seperation point, so the following analysis will breakdown the children to 0-3, 3-8, 8-18, and above 18

## Check the sample size
```{r}
Mapping_work %>% filter(SampleID %in% labels(DM_uu), SampleGroup=="Villagers") %>% group_by(Year, Ethnicity, Village, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID))) %>% ungroup() %>% mutate(Dummy = paste(Year, Age_grp2, sep = "-"), Year = NULL, Age_grp2 = NULL) %>% spread(key = Dummy, value = N_subject) 

Mapping_work %>% filter(SampleID %in% labels(DM_uu), SampleGroup=="Villagers") %>% group_by(Year, Expo, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID))) %>% spread(key = Age_grp2, value = N_subject) %>% kable 
```

## Figure \\label{PCoA_of_All}
** update on 6-26-2020: ** Given the discussion with CHOP, the PCOA should be calculated on each age group to have a better separation
```{r}
Indices = c("uu", "wu", "bc")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

beta_stats <- list()
for (YY in Yrs){
  #YY = "2015"
  for (BB in BodySites){
    # BB = "Feces"
    for (DD in Indices){
      #DD = "uu"
      Mf <- Mapping_work %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
      # distance matrix for YY, BB, and DD
      DM_tmp <- subset_dm_by_names(get(paste0("DM_", DD)), Mf$SampleID)
      # list of distanc matrix corresponding to each age group
      dms <- parallel::mclapply(c("Age_0-3", "Age_3-8", "Age_8-18", "Adults"), function(x){subset_dm_by_names(DM = DM_tmp, Sname = Mf %>% filter(Age_grp2==x) %>% pull(SampleID))}, mc.cores = nc)
      names(dms) <- c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")
      print(paste(YY, BB, DD))
      
      # Adonis test
      stats <- parallel::mclapply(dms, function(x){adonis2(x ~ Expo + Gender, by = "margin", data = Mf %>% filter(SampleID %in% attr(x, "Labels")))})
      print(stats)
      # beta dispersion
      stats_disp <- lapply(dms, function(x){betadisper(x, Mf[match(attr(x, "Labels"), Mf$SampleID), ]$Expo)})
      stats_disp_perm <- lapply(stats_disp, function(x){permutest(x)$tab$`Pr(>F)`[1]}) %>% do.call("rbind", .) %>% as.data.frame %>% rownames_to_column(var = "Age_grp2") %>% rename(p.disp = V1) %>% mutate(Factor = "Expo")
      
      stats_tbl <- lapply(seq_along(stats), function(x){(stats[[x]] %>% as.data.frame() %>% rownames_to_column(var = "Factor"))[1:2, c(1, 4:6)] %>% mutate(Age_grp2 = names(stats)[x] %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")))}) %>% do.call("rbind", .) %>% mutate(Body_Site = BB, Year = YY, Index = DD, R2 = round(R2, 3)) %>% merge(., stats_disp_perm, by = c("Age_grp2", "Factor"), all.x = T)
      
      stats_tbl2 <- stats_tbl %>% mutate(stat_label = paste0(Factor, ": p = ", `Pr(>F)`, "; R2 = ", R2), stat_label = ifelse(Factor=="Urban", paste0(stat_label, "; p.disp = ", p.disp), stat_label)) %>% group_by(Age_grp2) %>% summarise(stat_label = paste(stat_label[2], stat_label[1], sep = "\n")) 
      
      beta_stats[[paste0(YY, BB, DD)]] <- stats_tbl
      
      pcoa_tmp <- parallel::mclapply(dms, pcoa, mc.cores = nc)
      p_lst <- list()
      for (i in seq(length(pcoa_tmp))){
        varPC1 <- round(pcoa_tmp[[i]]$values$Relative_eig[1]*100, 2)
        varPC2 <- round(pcoa_tmp[[i]]$values$Relative_eig[2]*100, 2)
        varPC3 <- round(pcoa_tmp[[i]]$values$Relative_eig[3]*100, 2)
        plot_dt <- pcoa_tmp[[i]]$vectors[,1:3] %>% merge(., Mf, by.x = 0, by.y = 1, all.x = T)
        p_lst[[i]] <- ggplot(plot_dt, aes(x = Axis.1, y = Axis.2, color = Urban)) +
          geom_point(size = 1, aes(shape = Gender)) +
          stat_ellipse() +
          geom_label(data = stats_tbl2[i, ], x = -Inf, y = Inf, aes(label = stat_label), hjust = 0, vjust = 1, inherit.aes = F, size = 8/.pt) +
          facet_grid(.~Age_grp2) +
          theme_Publication(base_size = 10) + theme(aspect.ratio = 1, panel.background = element_rect(colour = "black"), panel.grid.major = element_line(colour="grey90",linetype="dashed", size=0.2), legend.direction = "vertical", strip.text = element_text(face="bold", color = "white", size = rel(1))) +
          labs(x = paste0("PC1: ", varPC1, "% explained variation"), y = paste0("PC2: ", varPC2, "% explained variation")) +
          scale_color_manual(breaks = names(Urban_color), values = Urban_color) +
          scale_shape_manual(values = c(16, 21))
      }
      p <- p_lst[[1]] + p_lst[[2]] + p_lst[[3]] + p_lst[[4]] + plot_layout(nrow = 1, guides = "collect") + plot_annotation(title = paste0("PCoA, ",YY, DD, BB), subtitle = "Rarefied at 10,000 reads per sample")
      print(p)
      set_panel_size(p = p, g = patchworkGrob(p), file = paste0("output/figures/pcoa_pc1v2_", DD, "_", BB, "_", YY, ".pdf"), width = unit(3, "in"), height = unit(3, "in"), useDingbats = F)
    }
  }
}

beta_stats_tbl <- do.call("rbind", beta_stats) %>% arrange(Index, Age_grp2, Factor, Body_Site, Year) %>% select(Index, Age_grp2, Factor, Body_Site, Year, everything()) 


clip <- pipe("pbcopy", "w")
write.table(beta_stats_tbl %>%  pivot_wider(id_cols = c(Index, Year, Body_Site, Factor), names_from = Age_grp2, values_from = c(R2, F, `Pr(>F)`, p.disp)), file = clip, row.names = F, quote = F, sep = "\t")
close(clip)

```

```{r}
Indices = c("uu", "wu", "bc")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

beta_stats <- list()
for (YY in Yrs){
    #YY = "2015"
    for (BB in BodySites){
        # BB = "Feces"
        for (DD in Indices){
            #DD = "uu"
            Mf <- Mapping_work2 %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
            # distance matrix for YY, BB, and DD
            DM_tmp <- subset_dm_by_names(get(paste0("DM_", DD)), Mf$SampleID)
            # Adonis test
            stats <- adonis2(DM_tmp ~ Urban + Age_num + Gender, by = "margin", data = Mf %>% filter(SampleID %in% attr(DM_tmp, "Labels")))
            print(stats)
            
            stats_tbl <- (stats %>% as.data.frame() %>% rownames_to_column(var = "Factor"))[1:3, c(1, 4:6)] %>% mutate(Body_Site = BB, Year = YY, Index = DD, R2 = round(R2, 3))
            
            beta_stats[[paste(YY, BB, DD, sep = "-")]] <- stats_tbl
        }
    }
}

beta_stats_tbl_3factor <- do.call("rbind", beta_stats)
beta_stats_tbl_3factor %>% knitr::kable()
```

## Figure \\label{PCoA_with_age}
** update on 07-02-2020: ** Given the discussion with Gloria, maybe a graph show of position of PC1 would be helpful

### Deicode
```{r}
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

for (YY in Yrs){
    #YY = "2015"
    p_lst <- list()
    p_lst2 <- list()
    for (BB in BodySites){
        # BB = "Feces"
        dat_name = paste(BB, YY, sep = "_")
        dat_ord = ord_qzas[[dat_name]]$data
        
        Mf <- Mapping_work2 %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
        
        varPC1 <- round(dat_ord$ProportionExplained$PC1*100, 2)
        varPC2 <- round(dat_ord$ProportionExplained$PC2*100, 2)
        varPC3 <- round(dat_ord$ProportionExplained$PC3*100, 2)
        
        # distance matrix for YY, BB, and DD
        DM_tmp <- subset_dm_by_names(dist_qzas[[dat_name]]$data, Mf$SampleID)
        
        
        # Adonis test
        stats <- adonis2(DM_tmp ~ Urban + Age_num + Gender, by = "margin", data = Mf %>% filter(SampleID %in% attr(DM_tmp, "Labels")))
        print(stats)
        
        stats_disp <- betadisper(DM_tmp, Mf[match(attr(DM_tmp, "Labels"), Mf$SampleID), ]$Urban) %>% permutest
        
        stats_tbl <- (stats %>% as.data.frame() %>% rownames_to_column(var = "Factor"))[1:3, c(1, 4:6)] %>% mutate(Body_Site = BB, Year = YY, Index = "Deicode", R2 = round(R2, 3))
        stats_tbl$p.disp[stats_tbl$Factor=="Urban"] = stats_disp$tab$`Pr(>F)`[1]
        ## renaming urban as exposure in the graph
        stats_tbl$Factor[stats_tbl$Factor=="Urban"] = "Exposure"
        
        stats_tbl2 <- stats_tbl %>% mutate(stat_label = paste0(Factor, ": p = ", `Pr(>F)`, "; R2 = ", R2), stat_label = ifelse(Factor=="Exposure", paste0(stat_label, "; p.disp = ", p.disp), stat_label)) %>% summarise(stat_label = paste(stat_label, collapse = "\n")) 
        
        
        
        plot_dt <- dat_ord$Vectors %>% merge(., Mf, by = 1)
        levels(plot_dt$Urban) <- c("Low", "Medium", "")

        
        p_lst[[BB]] <- ggplot(plot_dt, aes(x = Urban, y = PC1, fill = Age_grp2)) +
                geom_boxplot() + theme_bw(base_size = 8) + theme(aspect.ratio = 0.4) +
                geom_label(data = stats_tbl2, x = Inf, y = Inf, aes(label = stat_label), inherit.aes = F, size = 7/.pt, hjust = 1, vjust = 1) +
                scale_fill_manual(values = c("#f1eef6", "#bdc9e1", "#74a9cf", "#0570b0")) +
                scale_x_discrete(drop = F) +
                
                labs(x = "", y = paste0("PC1: ", varPC1, "% explained variation"), fill = "Age", title = BB) +
                coord_flip(clip = "off")
        p_lst2[[BB]] <- ggplot(plot_dt, aes(x = Urban, y = PC2, fill = Age_grp2)) +
                geom_boxplot() + theme_bw(base_size = 8) + theme(aspect.ratio = 0.4) +
                geom_label(data = stats_tbl2, x = Inf, y = Inf, aes(label = stat_label), inherit.aes = F, size = 7/.pt, hjust = 1, vjust = 1) +
                scale_fill_manual(values = c("#f1eef6", "#bdc9e1", "#74a9cf", "#0570b0")) +
                scale_x_discrete(drop = F) +
                
                labs(x = "", y = paste0("PC2: ", varPC2, "% explained variation"), fill = "Age", title = BB) +
                coord_flip(clip = "off")
    }
    wrap_plots(p_lst, ncol = 2, guides = "collect") + guide_area() + plot_annotation(tag_levels = "A", title = paste0("Deicode-", YY)) & theme(plot.margin = unit(c(3,3,3,3), "pt"))
    set_panel_size(p = last_plot(), g = patchworkGrob(last_plot()), file = paste0("output/figures/PC1_boxplot_deicode_all_", YY, ".pdf"), width = unit(3, "in"), height = unit(1.2, "in"), useDingbats = F)
    
    wrap_plots(p_lst2, ncol = 2, guides = "collect") + guide_area() + plot_annotation(tag_levels = "A", title = paste0("Deicode-", YY)) & theme(plot.margin = unit(c(3,3,3,3), "pt"))
    set_panel_size(p = last_plot(), g = patchworkGrob(last_plot()), file = paste0("output/figures/PC2_boxplot_deicode_all_", YY, ".pdf"), width = unit(3, "in"), height = unit(1.2, "in"), useDingbats = F)
    
}

```

### Other beta index
```{r}
Indices = c("uu", "wu", "bc")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

for (YY in Yrs){
    #YY = "2015"
    for (DD in Indices){
        #DD = "uu"
        p_lst <- list()
        for (BB in BodySites){
            # BB = "Feces"
            
            Mf <- Mapping_work2 %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
            
            # distance matrix for YY, BB, and DD
            DM_tmp <- subset_dm_by_names(get(paste0("DM_", DD)), Mf$SampleID)
            pcoa_tmp <- pcoa(DM_tmp)
            
             # Adonis test
            stats <- adonis2(DM_tmp ~ Urban + Age_num + Gender, by = "margin", data = Mf %>% filter(SampleID %in% attr(DM_tmp, "Labels")))
            print(stats)
            
            stats_disp <- betadisper(DM_tmp, Mf[match(attr(DM_tmp, "Labels"), Mf$SampleID), ]$Urban) %>% permutest
            
            stats_tbl <- (stats %>% as.data.frame() %>% rownames_to_column(var = "Factor"))[1:3, c(1, 4:6)] %>% mutate(Body_Site = BB, Year = YY, Index = DD, R2 = round(R2, 3))
            stats_tbl$p.disp[stats_tbl$Factor=="Urban"] = stats_disp$tab$`Pr(>F)`[1]
            ## renaming urban as exposure in the graph
            stats_tbl$Factor[stats_tbl$Factor=="Urban"] = "Exposure"
            
            stats_tbl2 <- stats_tbl %>% mutate(stat_label = paste0(Factor, ": p = ", `Pr(>F)`, "; R2 = ", R2), stat_label = ifelse(Factor=="Exposure", paste0(stat_label, "; p.disp = ", p.disp), stat_label)) %>% summarise(stat_label = paste(stat_label, collapse = "\n")) 
            
            varPC1 <- round(pcoa_tmp$values$Relative_eig[1]*100, 2)
            varPC2 <- round(pcoa_tmp$values$Relative_eig[2]*100, 2)
            varPC3 <- round(pcoa_tmp$values$Relative_eig[3]*100, 2)
            plot_dt <- pcoa_tmp$vectors[,1:3] %>% merge(., Mf, by.x = 0, by.y = 1, all.x = T)
            levels(plot_dt$Urban) <- c("Low", "Medium", "")
            

            p_lst[[BB]] <- ggplot(plot_dt, aes(x = Urban, y = Axis.1, fill = Age_grp2)) +
                geom_boxplot() + theme_bw(base_size = 8) + theme(aspect.ratio = 0.4) +
                geom_label(data = stats_tbl2, x = Inf, y = Inf, aes(label = stat_label), inherit.aes = F, size = 7/.pt, hjust = 1, vjust = 1) +
                scale_fill_manual(values = c("#f1eef6", "#bdc9e1", "#74a9cf", "#0570b0")) +
                scale_x_discrete(drop = F) +
                
                labs(x = "", y = paste0("PC1: ", varPC1, "% explained variation"), fill = "Age", title = BB) +
                coord_flip(clip = "off")
            last_plot()
        }
        wrap_plots(p_lst, ncol = 2, guides = "collect") + guide_area() + plot_annotation(tag_levels = "A", title = paste(DD, YY, sep = "-")) & theme(plot.margin = unit(c(3,3,3,3), "pt"))
        set_panel_size(p = last_plot(), g = patchworkGrob(last_plot()), file = paste0("output/figures/PC1_boxplot_",DD,"_all_", YY, ".pdf"), width = unit(3, "in"), height = unit(1.2, "in"), useDingbats = F)
    }
}

```



## Figure \\label{dist_child_to_adults} Distances of microbial communities between children and adults from the same village across time.

Unweighted unifrac, weighted unifrac, and bray-curtis
```{r}
compare_c_to_a_kw <- function(cc, a, f = "Urban"){
  ## c is the dataset for children, and a is the dataset for adults
  c_lst <- split(cc, as.character(cc[, "Urban"]))
  ### remove the High since it was generated due to factorized Urban variable
  
  a_lst <- split(a, as.character(a[, "Urban"]))
  
  c_a_kw <- lapply(seq_along(c_lst), function(x){kruskal.test(list(c_lst[[x]]$distance, a_lst[[x]]$distance))$p.value})
  names(c_a_kw) <- names(c_lst)
  
  c_a_kw <- do.call("rbind", c_a_kw) %>% as.data.frame()
  colnames(c_a_kw) <- "p.value"
  return(c_a_kw)
  
}
Indices = c("uu", "wu", "bc")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")


for (DD in Indices){
  #DD = "uu"
  for (YY in Yrs){
    #YY = "2015"
    p_combined <- list()
    for (BB in BodySites){
      #BB = "Feces"
      ## Extract dataset at specific year, body sites
      dat <- Mapping_work %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
      ## Since the analysis is done to compare children to adults distances at each villages, the dataset need to be separated according to village
      vls <- split(dat, dat$Village)
      ## get the distance table by getting all children to adult distance and the adult to adult distance at each villages, the children to children distance is not needed here.
      
      Dists <- lapply(vls, function(x){subset_dm_by_names(get(paste0("DM_", DD)), x$SampleID) %>% compare_group_dist(., x, "SampleID", "Age_grp1") %>% filter(Group_comp != "Children_Children") %>% mutate(SampleID = ifelse(Group_comp=="Adults_Adults", item1, ifelse(get_group_name(item1, x, "SampleID", "Age_grp1")=="Children", item1, item2)), item1 = NULL, item2 = NULL)}) ### here it doesn't matter which samples were compared for adult-adult distance, so we can use sample id of either item 1 or item 2 as the sample id outputed, otherwise, output the children sample id
      
      
      ## the Dists outputed from previous step is a list of dataframes (1 dataframe per village) in which 3 columns: distance, group, sample id, here the list is combined into a dataframe
      Dists <- do.call("rbind", Dists)
      ## split the dataframe by either adult-adult comparison, or adult-children comparison
      Dists2 <- Dists %>% split(Dists$Group_comp)
      ## average the distance between a child to different adults within the same village, this average will be compared in the statistics
      Dists2$Adults_Children <- Dists2$Adults_Children %>% group_by(Group_comp, SampleID) %>% summarise(distance = mean(distance)) %>% ungroup
      
      Dists2 <- do.call("rbind", Dists2)
      
      ## add the metadata to the above dataframe
      dat2 <- merge(Dists2, dat %>% select(SampleID, Age_num, Age_grp1, Ethnicity, Village, Urban), by = "SampleID", all.x = T)
      
      ## summarize the children results
      dat2_c <- dat2 %>% filter(Age_grp1=="Children")
      dat2_c_sum <- dat2_c %>% group_by(Urban, Age_num) %>% summarise(MCB = mean_cl_boot(distance) %>% paste(collapse = "_")) %>% separate(MCB, c("Mean", "Lwr", "Uwr"), sep = "_", convert = T) %>% ungroup
      ### stats on the children results, a non-parameter comparison between exposure levels will be done on a age-interval of 2 yrs, the p value will be adjusted using fdr (q<0.1)
      dat2_c_for_stat <- dat2_c %>% mutate(age_comp_grps = floor(Age_num/2)*2+0.5)
      dat2_c_for_stat_kw <- dat2_c_for_stat %>% group_by(age_comp_grps) %>% filter(length(unique(Urban))>1) %>% kruskal_test(distance ~ Urban) %>% merge(.,  dat2_c_for_stat %>% group_by(age_comp_grps, Urban) %>% summarise(Mean = mean(distance)) %>% filter(Mean==max(Mean)), by = 1) %>% ungroup() %>% mutate(fdr.q = p.adjust(p, method = "fdr") %>% round(3)) %>% filter(fdr.q<=0.1)
      
      ## separate the adults-adults distance
      dat2_a <- dat2 %>% filter(Age_grp1=="Adults")
      
      ## stats comparing children to adults
      dat2_c_to_a_stats <- lapply(split(dat2_c_for_stat, dat2_c_for_stat$age_comp_grps), function(x){compare_c_to_a_kw(x, dat2_a) %>% rownames_to_column(var = "Urban") %>% mutate(age_comp_grps = x$age_comp_grps[1], note = "Child_to_adults")}) %>% do.call("rbind", .) %>% ungroup() %>% mutate(fdr.q = p.adjust(p.value, method = "fdr") %>% round(3)) %>% filter(fdr.q<=0.1)
      
      ## plot the children-adult distance
      p1 <- ggplot(dat2_c, aes(x = Age_num, y = distance, color = Urban)) +
        geom_point(size = 1) +
        geom_ribbon(data = dat2_c_sum, aes(x = Age_num, ymin = Lwr, ymax = Uwr, fill = Urban), alpha = 0.5, inherit.aes = F) +
        geom_vline(xintercept = c(seq(1, 17, 2) + 0.5), linetype = "dotted") +
        stat_summary(geom = "line", fun.y = "mean") +
        theme_Publication(base_size = 10) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        scale_color_manual(breaks = names(Urban_color), values = Urban_color) +
        scale_fill_manual(breaks = names(Urban_color), values = Urban_color) +
        scale_x_continuous(breaks = seq(0, 18, 1), labels = sapply(seq(0,18,1), function(x){if(x%%2==0){""}else{x}})) +
        labs(x = "Age (Years)", y = paste(DD, "distance"), subtitle = "Children to adults", fill = "Exposure", color = "Exposure") + 
        coord_cartesian(clip = "off")
      
      
      ### check if any significance at each age interval
      #### following parameters are used to set the position of the stats on graph
      unit_y = (range(dat2_c$distance)[2]-range(dat2_c$distance)[1])/4
      y_c_comp = round(max(dat2_c$distance) + unit_y, 2)
      y_c_to_a_l = round(unit_y/3 + y_c_comp, 2)
      y_c_to_a_m = round(1.5 * unit_y/3 + y_c_to_a_l, 2)
      x_pos = max(dat2_c_for_stat_kw$age_comp_grps, dat2_c_to_a_stats$age_comp_grps) + 1
      if (nrow(dat2_c_for_stat_kw)>0){
        ## plot significant fdr.q value, the color indicated which groups is higher
        p1 <- p1 + geom_text(data = dat2_c_for_stat_kw, aes(x = age_comp_grps, y = y_c_comp, label = fdr.q, color = Urban), size = 7/.pt, show.legend = F) + annotate("label", x = x_pos, y = y_c_comp, size  = 7/.pt, label = "Low to Medium", fill = "white", hjust = 0)
      }
      if (nrow(dat2_c_to_a_stats)>0){
        dat2_c_to_a_stats <- dat2_c_to_a_stats %>% mutate(stat_y2 = ifelse(Urban=="Low", y_c_to_a_l, y_c_to_a_m))
        p1 <- p1 + geom_text(data = dat2_c_to_a_stats, aes(x = age_comp_grps, y = stat_y2, label = fdr.q, color = Urban), size = 7/.pt, show.legend = F) + annotate("label", x = x_pos, y = y_c_to_a_l, size  = 7/.pt, label = "Low to Adults", fill = "white", hjust = 0) + annotate("label", x = x_pos, y = y_c_to_a_m, size  = 7/.pt, label = "Medium to Adults", fill = "white", hjust = 0)
      }
      print(p1)
      
      p2 <- ggplot(dat2_a, aes(x = Urban, y = distance, color = Urban)) +
        geom_boxplot(show.legend = F) +
        theme_Publication(base_size = 10) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        scale_color_manual(breaks = names(Urban_color), values = Urban_color) +
        labs(x = "Exposure", y = paste(DD, "distance"), subtitle = "Within group distance among adults")
      
      print(p2)
      
      # 
      p_ranges_y <- c(ggplot_build(p1)$layout$panel_scales_y[[1]]$range$range, ggplot_build(p2)$layout$panel_scales_y[[1]]$range$range)
      
      p_combined[[BB]] <- p1 + p2 + plot_layout(guides = "collect", widths = c(3,1)) & scale_y_continuous(breaks = seq(round(min(p_ranges_y), 1), round(max(p_ranges_y), 1), 0.1), limits = c(min(p_ranges_y), max(p_ranges_y))) 
      p_combined[[BB]][[1]] <- p_combined[[BB]][[1]] + labs(tag = BB) + theme(plot.tag.position = c(0, 1), plot.tag = element_text(size = 11, face = "bold", hjust = 0, vjust = 0, margin = ggplot2::margin(b = 5)))
      print(p_combined[[BB]])
      set_panel_size(p = p_combined[[BB]], g = patchworkGrob(p_combined[[BB]]), file = paste0("output/figures/Distance_Children_", DD, "_", YY, "_", BB, ".pdf"), family = "Arial", width = unit(c(3, 1), "in"), height = unit(1.5, "in"), useDingbats = F)
      
    }
  } 
}

```

deicode
```{r}
compare_c_to_a_kw <- function(cc, a, f = "Urban"){
  ## c is the dataset for children, and a is the dataset for adults
  c_lst <- split(cc, as.character(cc[, "Urban"]))
  ### remove the High since it was generated due to factorized Urban variable
  
  a_lst <- split(a, as.character(a[, "Urban"]))
  
  c_a_kw <- lapply(seq_along(c_lst), function(x){kruskal.test(list(c_lst[[x]]$distance, a_lst[[x]]$distance))$p.value})
  names(c_a_kw) <- names(c_lst)
  
  c_a_kw <- do.call("rbind", c_a_kw) %>% as.data.frame()
  colnames(c_a_kw) <- "p.value"
  return(c_a_kw)
  
}
Indices = c("deicode")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")



for (YY in Yrs){
  #YY = "2015"
  p_combined <- list()
  for (BB in BodySites){
    #BB = "Feces"
    ## Extract dataset at specific year, body sites
    dat <- Mapping_work %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
    ## Since the analysis is done to compare children to adults distances at each villages, the dataset need to be separated according to village
    vls <- split(dat, dat$Village)
    ## get the distance table by getting all children to adult distance and the adult to adult distance at each villages, the children to children distance is not needed here.
    
    Dists <- lapply(vls, function(x){subset_dm_by_names(DM_deicode[[paste0(BB, "_", YY)]], x$SampleID) %>% compare_group_dist(., x, "SampleID", "Age_grp1") %>% filter(Group_comp != "Children_Children") %>% mutate(SampleID = ifelse(Group_comp=="Adults_Adults", item1, ifelse(get_group_name(item1, x, "SampleID", "Age_grp1")=="Children", item1, item2)), item1 = NULL, item2 = NULL)}) ### here it doesn't matter which samples were compared for adult-adult distance, so we can use sample id of either item 1 or item 2 as the sample id outputed, otherwise, output the children sample id
    
    
    ## the Dists outputed from previous step is a list of dataframes (1 dataframe per village) in which 3 columns: distance, group, sample id, here the list is combined into a dataframe
    Dists <- do.call("rbind", Dists)
    ## split the dataframe by either adult-adult comparison, or adult-children comparison
    Dists2 <- Dists %>% split(Dists$Group_comp)
    ## average the distance between a child to different adults within the same village, this average will be compared in the statistics
    Dists2$Adults_Children <- Dists2$Adults_Children %>% group_by(Group_comp, SampleID) %>% summarise(distance = mean(distance)) %>% ungroup()
    
    Dists2 <- do.call("rbind", Dists2)
    
    ## add the metadata to the above dataframe
    dat2 <- merge(Dists2, dat %>% select(SampleID, Age_num, Age_grp1, Ethnicity, Village, Urban), by = "SampleID", all.x = T)
    
    ## summarize the children results
    dat2_c <- dat2 %>% filter(Age_grp1=="Children")
    dat2_c_sum <- dat2_c %>% group_by(Urban, Age_num) %>% summarise(MCB = mean_cl_boot(distance) %>% paste(collapse = "_")) %>% separate(MCB, c("Mean", "Lwr", "Uwr"), sep = "_", convert = T) %>% ungroup
    ### stats on the children results, a non-parameter comparison between exposure levels will be done on a age-interval of 2 yrs, the p value will be adjusted using fdr (q<0.1)
    dat2_c_for_stat <- dat2_c %>% mutate(age_comp_grps = floor(Age_num/2)*2+0.5)
    dat2_c_for_stat_kw <- dat2_c_for_stat %>% group_by(age_comp_grps) %>% filter(length(unique(Urban))>1) %>% kruskal_test(distance ~ Urban) %>% merge(.,  dat2_c_for_stat %>% group_by(age_comp_grps, Urban) %>% summarise(Mean = mean(distance)) %>% filter(Mean==max(Mean)), by = 1) %>% ungroup() %>% mutate(fdr.q = p.adjust(p, method = "fdr") %>% round(3)) %>% filter(fdr.q<=0.1)
    
    ## separate the adults-adults distance
    dat2_a <- dat2 %>% filter(Age_grp1=="Adults")
    
    ## stats comparing children to adults
    dat2_c_to_a_stats <- lapply(split(dat2_c_for_stat, dat2_c_for_stat$age_comp_grps), function(x){compare_c_to_a_kw(x, dat2_a) %>% rownames_to_column(var = "Urban") %>% mutate(age_comp_grps = x$age_comp_grps[1], note = "Child_to_adults")}) %>% do.call("rbind", .) %>% ungroup() %>% mutate(fdr.q = p.adjust(p.value, method = "fdr") %>% round(3)) %>% filter(fdr.q<=0.1)
    
    ## plot the children-adult distance
    p1 <- ggplot(dat2_c, aes(x = Age_num, y = distance, color = Urban)) +
      geom_point(size = 1) +
      geom_ribbon(data = dat2_c_sum, aes(x = Age_num, ymin = Lwr, ymax = Uwr, fill = Urban), alpha = 0.5, inherit.aes = F) +
      geom_vline(xintercept = c(seq(1, 17, 2) + 0.5), linetype = "dotted") +
      stat_summary(geom = "line", fun = "mean") +
      theme_Publication(base_size = 10) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      scale_color_manual(breaks = names(Urban_color), values = Urban_color) +
      scale_fill_manual(breaks = names(Urban_color), values = Urban_color) +
      scale_x_continuous(breaks = seq(0, 18, 1), labels = sapply(seq(0,18,1), function(x){if(x%%2==0){""}else{x}})) +
      labs(x = "Age (Years)", y = "Deicode Distance", subtitle = "Children to adults", fill = "Exposure", color = "Exposure") + 
      coord_cartesian(clip = "off")
    
    
    ### check if any significance at each age interval
    #### following parameters are used to set the position of the stats on graph
    unit_y = (range(dat2_c$distance)[2]-range(dat2_c$distance)[1])/4
    y_c_comp = round(max(dat2_c$distance) + unit_y, 2)
    y_c_to_a_l = round(unit_y/3 + y_c_comp, 2)
    y_c_to_a_m = round(1.5 * unit_y/3 + y_c_to_a_l, 2)
    x_pos = max(dat2_c_for_stat_kw$age_comp_grps, dat2_c_to_a_stats$age_comp_grps) + 1
    if (nrow(dat2_c_for_stat_kw)>0){
      ## plot significant fdr.q value, the color indicated which groups is higher
      p1 <- p1 + geom_text(data = dat2_c_for_stat_kw, aes(x = age_comp_grps, y = y_c_comp, label = fdr.q, color = Urban), size = 7/.pt, show.legend = F) + annotate("label", x = x_pos, y = y_c_comp, size  = 7/.pt, label = "Low to Medium", fill = "white", hjust = 0)
    }
    if (nrow(dat2_c_to_a_stats)>0){
      dat2_c_to_a_stats <- dat2_c_to_a_stats %>% mutate(stat_y2 = ifelse(Urban=="Low", y_c_to_a_l, y_c_to_a_m))
      p1 <- p1 + geom_text(data = dat2_c_to_a_stats, aes(x = age_comp_grps, y = stat_y2, label = fdr.q, color = Urban), size = 7/.pt, show.legend = F) + annotate("label", x = x_pos, y = y_c_to_a_l, size  = 7/.pt, label = "Low to Adults", fill = "white", hjust = 0) + annotate("label", x = x_pos, y = y_c_to_a_m, size  = 7/.pt, label = "Medium to Adults", fill = "white", hjust = 0)
    }
    print(p1)
    
    p2 <- ggplot(dat2_a, aes(x = Urban, y = distance, color = Urban)) +
      geom_boxplot(show.legend = F) +
      theme_Publication(base_size = 10) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      scale_color_manual(breaks = names(Urban_color), values = Urban_color) +
      labs(x = "Exposure", y = "Deicode distance", subtitle = "Within group distance among adults")
    
    print(p2)
    
    # 
    p_ranges_y <- c(ggplot_build(p1)$layout$panel_scales_y[[1]]$range$range, ggplot_build(p2)$layout$panel_scales_y[[1]]$range$range)
    
    p_combined[[BB]] <- p1 + p2 + plot_layout(guides = "collect", widths = c(3,1)) & scale_y_continuous(limits = c(min(p_ranges_y), max(p_ranges_y))) 
    p_combined[[BB]][[1]] <- p_combined[[BB]][[1]] + labs(tag = BB) + theme(plot.tag.position = c(0, 1), plot.tag = element_text(size = 11, face = "bold", hjust = 0, vjust = 0, margin = ggplot2::margin(b = 5)))
    print(p_combined[[BB]])
    set_panel_size(p = p_combined[[BB]], g = patchworkGrob(p_combined[[BB]]), file = paste0("output/figures/Distance_Children_Deicode", "_", YY, "_", BB, ".pdf"), family = "Arial", width = unit(c(3, 1), "in"), height = unit(1.5, "in"), useDingbats = F)
    
  }
} 


```
