---
title: "Human body morphometry"
date: "`r Sys.Date()`"
output:
  html_document:
      code_folding: hide

      df_print: kable
editor_options:
  chunk_output_type: inline
---

```{r setup, message=FALSE, warning=FALSE}
library(knitr)
options(max.print="999")
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, prompt=FALSE, comment = NA, message = FALSE)
opts_knit$set(width=75)

# Loading library here
library(tidyverse)
library(extrafont)
library(ggpubr)

# set path
path <- switch(Sys.info()[['sysname']], Windows = "C:/Users/jinch/", Darwin = "~/")
opts_knit$set(root.dir = paste0(path, "Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/91-Analysis-Main"))
source(paste0(path, "Dropbox/42-JCHWANG-RUTGERS/Projects-Rutgers/Src/utils.R"))
setwd(paste0(path, "Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/91-Analysis-Main"))
```

**Setup color index**
```{r}
# Urban level
Urban_color <- c("#377EB8", "#E41A1C", "#4DAF4A")
names(Urban_color) <- c("Low", "Medium", "High")
print(Urban_color)
# Ethnicity
Ethnicity_color <- gg_color_hue(3)
names(Ethnicity_color) <- c("SANEMA", "YEKWANA", "Visitors")
print(Ethnicity_color)

Sanemas <- c("Chajuranha", "Mosenahanha", "Kuyuwininha", "Shianana-Jiyakwanha", "Washudihanha", "Sudukuma")
Yekwana <- c("Kanarakuni", "Fiyakwanha")
```


** Mapping files**
Mapping files that were used in this analysis are:

`Mapping_Bolivar_Villagers-20181108.txt` (md5): `47c35f47f6e9ce45e0bf9c052525eee2`

```{r}
# Using revised mapping file
Mapping_VL <- read.delim("../Mapping_Bolivar_Villagers-20200416-jw.txt", stringsAsFactors = F)
Mapping_VL[is.na(Mapping_VL)] <- "NA"

## remove village "Kadanskaduinha" becaue very few subjects, and those without age information
Mapping_MS_VL <- Mapping_VL %>% filter(Village != "Kadansakaduinha", Age!="NA")
Mapping_MS_VL$Ethnicity[which(Mapping_MS_VL$Ethnicity=="Yekwana")] <- "YEKWANA"
```


## Analysis
```{r}
Mapping_work <- Mapping_MS_VL %>% mutate(Exposure = ifelse(Ethnicity=="SANEMA", "Low", ifelse(Village=="Fiyakwanha", "Low", ifelse(SampleGroup=="Visitors", "High", "Medium"))) %>% factor(levels = c("Low", "Medium", "High")), Age_num = as.numeric(Age), Age_num = ifelse(is.na(Age_num), as.numeric(gsub(pattern = "_months", replacement = "", Age, ignore.case = T))/12, Age_num), Age_grp1 = ifelse(Age_num>=18, "Adults", "Children"))
```

### Check the sample size
```{r}
Mapping_work %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Subject_ID_unique = paste(Village, House_ID, Family_ID, Subject_ID, sep = "-"))%>% group_by(Year, Ethnicity, Village, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID_unique))) %>% ungroup() %>% mutate(Dummy = paste(Year, Age_grp2, sep = "-"), Year = NULL, Age_grp2 = NULL) %>% spread(key = Dummy, value = N_subject)

# Check data availability for weight, height, BMI
## weight
Mapping_work %>% filter(Weight!="NA") %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Subject_ID_unique = paste(Village, House_ID, Family_ID, Subject_ID, sep = "-"))%>% group_by(Year, Ethnicity, Village, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID_unique))) %>% ungroup() %>% mutate(Dummy = paste(Year, Age_grp2, sep = "-"), Year = NULL, Age_grp2 = NULL) %>% spread(key = Dummy, value = N_subject)

## height
Mapping_work %>% filter(Height!="NA") %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Subject_ID_unique = paste(Village, House_ID, Family_ID, Subject_ID, sep = "-"))%>% group_by(Year, Ethnicity, Village, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID_unique))) %>% ungroup() %>% mutate(Dummy = paste(Year, Age_grp2, sep = "-"), Year = NULL, Age_grp2 = NULL) %>% spread(key = Dummy, value = N_subject)

## BMI
Mapping_work %>% filter(BMI!="NA") %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Subject_ID_unique = paste(Village, House_ID, Family_ID, Subject_ID, sep = "-"))%>% group_by(Year, Ethnicity, Village, Age_grp2) %>% summarise(N_subject = length(unique(Subject_ID_unique))) %>% ungroup() %>% mutate(Dummy = paste(Year, Age_grp2, sep = "-"), Year = NULL, Age_grp2 = NULL) %>% spread(key = Dummy, value = N_subject)

```

Weight, Height, and BMI were avaialable for all villagers from Kanarakuni in 2015 and most villagers in Kanarakuni in 2016; none for Sanema villagers in 2015, some for sanema villagers and Fiyakwanha in 2016. So unless the analysis is based on Kanarakuni only, it is not possible to provide the information.

### Weight, height and BMI in subset populations in 2016
```{r}
dat  <- Mapping_work %>% filter(Year==2016, BMI!="NA", Weight!="NA", Height!="NA") %>% mutate_at(vars(BMI, Weight, Height), .funs = as.numeric)

ggplot(dat, aes(x = Age_num, y = Weight, color = Exposure)) +
  geom_point()

  
ggplot(sdad) +

```



