---
title: "ANCOM analysis"
date: "`r Sys.Date()`"
output:
  html_document:
      theme: simplex
      code_folding: hide
      toc: true
      toc_depth: 2
      toc_float: true
editor_options:
  chunk_output_type: console
---

## Preparation
1. Setup
```{r setup, message=FALSE, warning=FALSE}
# Loading library here
library(conflicted)
library(knitr)
library(tidyverse)
library(extrafont)
#library(biomformat)
#library(pheatmap)
library(rstatix)
library(data.tree)
library(ggtree)
library(treeio)

conflict_prefer("filter", "dplyr")

options(max.print="75")
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, prompt=FALSE, tidy = TRUE, comment = NA, message = FALSE)
opts_knit$set(width=75)

# set path
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source("~/Dropbox/42-JCHWANG-RUTGERS/Projects-Rutgers/Src/utils.R")
setwd(rprojroot::find_rstudio_root_file())
#devtools::source_url("https://raw.githubusercontent.com/sidhujyatha/ANCOM/master/ANCOM_updated_code_06_06_2020.R", sha1 = "c38079a8387079d5218605142578c8d7767925d5") # this is an updated version of Ancom2 code with corrections on the formula
# do not need to load ancom code since already analyzed.
# source("~/Dropbox/42-JCHWANG-RUTGERS/Projects-Rutgers/Src/ANCOM_updated_code_06_06_2020.R")

# number of cores/thread to use
nc <- switch(Sys.info()[['sysname']], Windows = 1, Darwin = 4, Linux = 12)

sessionInfo()
```

2. Color index
```{r}
# Urban level
Urban_color <- c("#66c2a5", "#fc8d62", "#8da0cb")
names(Urban_color) <- c("Low", "Medium", "High")
print(Urban_color)
# Ethnicity
Ethnicity_color <- gg_color_hue(3)
names(Ethnicity_color) <- c("SANEMA", "YEKWANA", "Visitors")
```

3. Mapping file
```{r}
# Mapping
## This mapping files include all villagers and baseline of visitors of all body sites
load("output/Mapping_MS_Urban.Rdata")
Mapping_work <- Mapping_MS_Urban %>% mutate(Urban = ifelse(Ethnicity=="SANEMA", "Low", ifelse(Village=="Fiyakwanha", "Low", ifelse(SampleGroup=="Visitors", "High", "Medium"))) %>% factor(levels = c("Low", "Medium", "High")), Age_num = as.numeric(Age), Age_num = ifelse(is.na(Age_num), as.numeric(gsub(pattern = "_months", replacement = "", Age, ignore.case = T))/12, Age_num), Age_grp1 = ifelse(Age_num>=18, "Adults", "Children"))
Mapping_work$Age_num[Mapping_work$Age=="4_Days"] <- 0
Mapping_work$Age_num[Mapping_work$Age=="1_day"] <- 0
Mapping_work$Age_grp1[Mapping_work$Age_num==0] <- "Children"

Mapping_work2 <- Mapping_work %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<=18 ~ "Age_8-18", T ~ "Adults") %>% factor(levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")))
```

4. Feature table (do not need to run)
This below section are used to generate taxa glom feature table
```{r, eval = FALSE}
# This is the rarefied biom table
load("output/FT_rarefied_full_11142019.Rdata")
## select only those appeared in the mapping files, and filter "OTU" that has 0 in all selected samples
Work_biom <- biom %>% rownames_to_column(var = "OTU_ID") %>% dplyr::select(OTU_ID, any_of(Mapping_work$SampleID)) %>% filter(rowSums(.[,-1])!=0)

# Taxa table
taxa <- read.delim("~/Dropbox/41-JCHWANG-NYU/Projects/10-Bolivar2018/3-16SDataProcessing/5-Taxonomy/export/silva132/data/taxonomy.tsv", stringsAsFactors = F) %>%  mutate(Taxon =  gsub(pattern = "__", replacement = "_", Taxon)) %>% separate(Taxon, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") # 7b9c4a9d-2c25-4ea9-a091-0770d3c3b39b

taxa$Domain[is.na(taxa$Domain)] <- "D_0_"
taxa$Phylum[is.na(taxa$Phylum)] <- "D_1_"
taxa$Class[is.na(taxa$Class)] <- "D_2_"
taxa$Order[is.na(taxa$Order)] <- "D_3_"
taxa$Family[is.na(taxa$Family)] <- "D_4_"
taxa$Genus[is.na(taxa$Genus)] <- "D_5_"
taxa$Species[is.na(taxa$Species)] <- "D_6_"
Work_taxa <- taxa %>% filter(Feature.ID %in% Work_biom$OTU_ID)

# Biom table with taxa
ft_w_taxa <- merge(Work_taxa[, -9], Work_biom, by = 1)

# Create a table with summation at levels from Domain to Genus
ft_glom <- list()
for (tt in c("Genus", "Family", "Order", "Class", "Phylum", "Domain")){
    ft_glom[[tt]] <- ft_w_taxa %>% group_by_at(vars("Domain":tt)) %>% summarise(across(starts_with("X"), sum, na.rm = T)) %>% ungroup() %>% unite("Taxa", "Domain":tt, sep = "|", remove = F)
}

# this datafile is combining the above list into a single table.
ft_glom_df <- data.table::rbindlist(ft_glom, use.names = T, fill = T, idcol = "taxa.glom")
save(ft_glom, file = "output/FT_rarefied_w_taxa_glom.Rdata")
save(ft_glom_df, file = "output/FT_rarefied_w_taxa_glom_df.Rdata")
```

## Differential Analysis
```{r}
load("output/FT_rarefied_w_taxa_glom_df.Rdata")
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

```
### Ancom
```{r functions, eval=FALSE}
ancom_wrap <- function(featuretable, metadata, ...){
    ft <- featuretable # feature table is in a format that a column of taxa with additional columns of counts from each samples
    mt <- metadata
    
    # filter feature table to remove features that appeared in less than 10% of samples
    sample_names <- colnames(ft)[-1]
    n_samples <- length(sample_names)
    ft_flt <- ft[rowSums(ft[, -1]>0)>=0.1*n_samples, ]
    
    ## create ancom format feature table
    ft_ancom <- ft_flt[, -1] %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Sample.ID")
    colnames(ft_ancom) <- c("Sample.ID", ft_flt$Taxa) %>% make.names()
    
    
    # filter metadata
    mt_ancom <- mt %>% filter(SampleID %in% ft_ancom$Sample.ID) %>% rename(Sample.ID = SampleID)
    
    ancom_mod <- ANCOM.main(OTUdat = ft_ancom, Vardat = mt_ancom, ...)
    return(ancom_mod)
    
}
```



#### per year, per sites
```{r ancom_run, eval=FALSE}
Ancom_mf_lst_2 <- Mapping_work2 %>% filter(SampleGroup=="Villagers", Body_Site %in% BodySites) %>% group_by(Year, Body_Site) %>% group_split()

Ancom_rsl_2 <- parallel::mclapply(Ancom_mf_lst_2, function(x){
    list(x[1, c("Year", "Body_Site")], ancom_wrap(ft_glom$Genus %>% select(Taxa, any_of(x$SampleID)), x, adjusted = T, main.var = "Urban", adj.formula = "Gender+Age_num", repeat.var = NULL, longitudinal = F, random.formula = NULL, multcorr = 2, sig = 0.05, prev.cut = 1))
    }, mc.cores = nc)

# Ancom_rsl_3 <- parallel::mclapply(Ancom_mf_lst_2, function(x){
#     list(x[1, c("Year", "Body_Site")], ancom_wrap(ft_glom$Family %>% select(Taxa, any_of(x$SampleID)), x, adjusted = T, main.var = "Urban", adj.formula = "Gender+Age_num", repeat.var = NULL, longitudinal = F, random.formula = NULL, multcorr = 2, sig = 0.05, prev.cut = 1))
# }, mc.cores = 20)
# 
# Ancom_rsl_4 <- parallel::mclapply(Ancom_mf_lst_2, function(x){
#     list(x[1, c("Year", "Body_Site")], ancom_wrap(ft_glom$Order %>% select(Taxa, any_of(x$SampleID)), x, adjusted = T, main.var = "Urban", adj.formula = "Gender+Age_num", repeat.var = NULL, longitudinal = F, random.formula = NULL, multcorr = 2, sig = 0.05, prev.cut = 1))
# }, mc.cores = 20)
# 
# Ancom_rsl_5 <- parallel::mclapply(Ancom_mf_lst_2, function(x){
#     list(x[1, c("Year", "Body_Site")], ancom_wrap(ft_glom$Class %>% select(Taxa, any_of(x$SampleID)), x, adjusted = T, main.var = "Urban", adj.formula = "Gender+Age_num", repeat.var = NULL, longitudinal = F, random.formula = NULL, multcorr = 2, sig = 0.05, prev.cut = 1))
# }, mc.cores = 20)
# 
# Ancom_rsl_6 <- parallel::mclapply(Ancom_mf_lst_2, function(x){
#     list(x[1, c("Year", "Body_Site")], ancom_wrap(ft_glom$Phylum %>% select(Taxa, any_of(x$SampleID)), x, adjusted = T, main.var = "Urban", adj.formula = "Gender+Age_num", repeat.var = NULL, longitudinal = F, random.formula = NULL, multcorr = 2, sig = 0.05, prev.cut = 1))
# }, mc.cores = 20)
    

save(Ancom_rsl_2, file = "output/Ancom_rsl_g_YYBB.Rdata")
# save(Ancom_rsl_3, file = "output/Ancom_rsl_f_YYBB.Rdata")
# save(Ancom_rsl_4, file = "output/Ancom_rsl_o_YYBB.Rdata")
# save(Ancom_rsl_5, file = "output/Ancom_rsl_c_YYBB.Rdata")
# save(Ancom_rsl_6, file = "output/Ancom_rsl_p_YYBB.Rdata")
```

```{r ancom_summarizing}
load("output/Ancom_rsl_g_YYBB.Rdata")
load("output/Ancom_rsl_f_YYBB.Rdata")
load("output/Ancom_rsl_o_YYBB.Rdata")
load("output/Ancom_rsl_c_YYBB.Rdata")
load("output/Ancom_rsl_p_YYBB.Rdata")

# summarize ancom results
Ancom_rsl_2_tbl <- lapply(Ancom_rsl_2, function(x){x[[2]]$W.taxa %>% mutate(Year = as.character(x[[1]][1,1]), Body_Site = as.character(x[[1]][1,2]))}) %>% do.call("rbind", .) %>% 
  filter(!grepl(pattern = "D_1_\\.D", otu.names)) # remove those genera with no phylum or domain info
# for family, order, class, and phylum, only those LT0.9 selected as significant
Ancom_rsl_3_tbl <- lapply(Ancom_rsl_3, function(x){x[[2]]$W.taxa %>% mutate(Year = as.character(x[[1]][1,1]), Body_Site = as.character(x[[1]][1,2]))}) %>% do.call("rbind", .) %>% filter(!grepl(pattern = "D_1_\\.D", otu.names))
Ancom_rsl_4_tbl <- lapply(Ancom_rsl_4, function(x){x[[2]]$W.taxa %>% mutate(Year = as.character(x[[1]][1,1]), Body_Site = as.character(x[[1]][1,2]))}) %>% do.call("rbind", .) %>% filter(!grepl(pattern = "D_1_\\.D", otu.names)) 
Ancom_rsl_5_tbl <- lapply(Ancom_rsl_5, function(x){x[[2]]$W.taxa %>% mutate(Year = as.character(x[[1]][1,1]), Body_Site = as.character(x[[1]][1,2]))}) %>% do.call("rbind", .) %>% filter(!grepl(pattern = "D_1_\\.D", otu.names)) 
Ancom_rsl_6_tbl <- lapply(Ancom_rsl_6, function(x){x[[2]]$W.taxa %>% mutate(Year = as.character(x[[1]][1,1]), Body_Site = as.character(x[[1]][1,2]))}) %>% do.call("rbind", .) %>% filter(!grepl(pattern = "D_1_\\.D", otu.names)) 

All_genus_tbl <- Ancom_rsl_2_tbl %>% distinct(otu.names) %>% mutate(otu.names = as.character(otu.names), V1 = gsub(pattern = "\\.D", replacement = "/D", otu.names)) %>% separate(V1, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "/") %>% mutate(Genus_index = paste0("G", seq(n())))
```

##### Check significant changed genus 

```{r eval=FALSE}
sig_cutoff <- integer()
for (YY in Yrs){
  # YY = 2015
  for (BB in BodySites){
    # BB = "Feces"
    ## metadata
    Mf <- Mapping_work2 %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
    ## Ancom results and catogrize the species according to W/effect size
    AR <- Ancom_rsl_2_tbl %>% filter(Year==YY, Body_Site==BB) %>% mutate(otu.names = as.character(otu.names)) %>% arrange(desc(W_stat))
    ### check the distribution of W-stats, 
    p <- ggplot(AR, aes(x = W_stat)) +
      geom_histogram(aes(y = ..density..), bins = 20) +
      geom_density() +
      geom_vline(xintercept = nrow(AR)*seq(0.6, 0.9, 0.1)) + 
      theme_bw() +
      labs(title = paste(YY, BB, sep = "-"))
    print(p)
    
    table(AR$ES)
    sig_cutoff[paste(YY, BB, sep = "-")] <- readline(prompt = "Above which cut off deemed as significant? LT0.6(1), LT0.7(2), LT0.8(3), LT0.9(4)")
  }
}

```
```{r}
sig_cutoff = c(1, 2, 2, 3, 3, 1, 3, 1, 3, 3)
names(sig_cutoff) = c("2015-Feces", "2015-Mouth", "2015-Nose", "2015-Right_Arm", "2015-Right_Hand", "2016-Feces", "2016-Mouth", "2016-Nose", "2016-Right_Arm", "2016-Right_Hand")

```

The vector `sig_cutoff` stores the cutoff that arbitrarily determines above which level to be considered as significant LT0.6(1), LT0.7(2), LT0.8(3), LT0.9(4).

##### Create data tables
```{r}
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")

# mapping and taxa files for feature tables
fts_mt <- parallel::mclapply(cross2(BodySites, Yrs), function(x){
    list(Mf = Mapping_work2 %>% filter(Year==x[[2]], Body_Site==x[[1]], SampleGroup=="Villagers"),
         AR = Ancom_rsl_2_tbl %>% filter(Year==x[[2]], Body_Site==x[[1]]) %>% mutate(otu.names = as.character(otu.names)) %>% arrange(desc(W_stat)) ,
         AR_f = Ancom_rsl_3_tbl %>% filter(Year==x[[2]], Body_Site==x[[1]]) %>% mutate(otu.names = as.character(otu.names)),
         AR_o = Ancom_rsl_4_tbl %>% filter(Year==x[[2]], Body_Site==x[[1]]) %>% mutate(otu.names = as.character(otu.names)),
         AR_c = Ancom_rsl_5_tbl %>% filter(Year==x[[2]], Body_Site==x[[1]]) %>% mutate(otu.names = as.character(otu.names)),
         AR_p = Ancom_rsl_6_tbl %>% filter(Year==x[[2]], Body_Site==x[[1]]) %>% mutate(otu.names = as.character(otu.names)))}, mc.cores = nc)
# feature tables
fts <- parallel::mclapply(fts_mt, function(x){ft_glom_df %>% select(taxa.glom:Genus, any_of(x$Mf$SampleID)) %>% mutate(Taxa = make.names(Taxa)) %>% filter(Taxa %in% unlist(lapply(x[2:6], function(x){pull(x, var = "otu.names")})))}, mc.cores = nc)

```

##### check proportions of taxa by W_stat category **(need updates)**
```{r}
for (i in seq_along(fts)){
    dt_tmp <- fts[[i]] %>% gather(key = "SampleID", value = "Count", -1) %>% merge(., fts_mt[[i]][["Mf"]], by = "SampleID", all.x = T) %>% merge(., fts_mt[[i]][["AR"]] %>% select(otu.names, W_stat, ES), by.x = "Taxa", by.y = "otu.names") %>% mutate(Rel_Ab = Count/10000) %>%  arrange(Urban, Age_num)
    p <- ggplot(dt_tmp, aes(x = SampleID, y = Rel_Ab, fill = ES)) +
        geom_bar(stat = "identity") +
        facet_grid(.~Urban + Gender, scales = "free_x") +
        labs(title = paste(dt_tmp$Year[1], dt_tmp$Body_Site[1], sep = "-")) +
        theme_bw() + theme(axis.text.x = element_blank())
    print(p)
    
    p <- ggplot(dt_tmp %>% group_by(Urban, Gender, Age_grp2, Age_num, SampleID, ES) %>% summarise(Rel_Ab2 = sum(Rel_Ab)) , aes(x = Rel_Ab2, y = ES)) +
        geom_violin(scale = "count") +
        facet_grid(.~Urban) +
        theme_bw() +
        labs(x = "Relative Abundance", y = "W stats categories", title = paste(dt_tmp$Year[1], dt_tmp$Body_Site[1], sep = "-"))
    print(p)
}
```

##### making graphs
0. Calculate odds ratios
```{r}
lrs_b_grp <- list()
for (ss in seq_along(fts)){
    # ss = 3
    # prepare datasets
    ft_t <- fts[[ss]]
    Mf <- fts_mt[[ss]][["Mf"]] %>% filter(SampleID %in% colnames(ft_t))
    YY = Mf$Year[1]
    BB = Mf$Body_Site[1]
    
    # relative abundance
    ft_t_rel_ab <- ft_t %>% mutate(across(starts_with("X"), ~ .x/10000))
        
    # rel abundance in log scale
    ft_t_rel_ab_log10 <- ft_t_rel_ab %>% mutate(across(starts_with("X"), ~ ifelse(.x == 0, log10(0.0001), log10(.x))))
    
    # calcualate the log ratio of a taxa between medium and low exposure level
    lr_b_grp <- split(ft_t_rel_ab_log10, ft_t_rel_ab_log10$Taxa) %>% lapply(., function(x){10^((t(x[, -1]) %>% merge(., Mf, by.x = 0, by.y = 1) %>% lm(V1 ~ Urban + Age_grp2 + Gender, data = .))$coefficient[2]) %>% log2() }) %>% unlist(use.names = T)
    names(lr_b_grp) <- names(lr_b_grp) %>% gsub(pattern = "\\.UrbanMedium", replacement = "", x = .)
    lrs_b_grp[[ss]] <- lr_b_grp
}
```
1. Heatmap of significantly changed genus
```{r}
for (ss in seq_along(fts)){
    # ss = 3
    # prepare datasets
    ft_t <- fts[[ss]]
    Mf <- fts_mt[[ss]][["Mf"]] %>% filter(SampleID %in% colnames(ft_t))
    YY = Mf$Year[1]
    BB = Mf$Body_Site[1]
    lr_b_grp = lrs_b_grp[[ss]]
    AR_g <- fts_mt[[ss]]$AR
    # defining taxa that changed and unchanged based on sig_cutoff
    ind = paste(YY, BB, sep = "-")
    ES_cutoff <- seq(0.6, 0.9, 0.1)[sig_cutoff[ind]] * nrow(AR_g)
    
    # relative abundance
    ft_t_rel_ab <- ft_t %>% mutate(across(starts_with("X"), ~ .x/10000))
    
    # rel abundance in log scale
    ft_t_rel_ab_log10 <- ft_t_rel_ab %>% mutate(across(starts_with("X"), ~ ifelse(.x == 0, log10(0.0001), log10(.x))))
    
    # heatmap of significantly changed features
    ## dataframe for the column annotation
    df_col <- Mf %>% select(SampleID, Age_grp2, Urban, Gender) %>% mutate(Urban =  as.character(Urban)) %>% arrange(Urban, Age_grp2) %>% column_to_rownames(var = "SampleID")
    ## dataframe for the row(taxa) annotation, only significantly modified genera included
    df_row <- data.frame(LR = lr_b_grp[AR_g %>% filter(W_stat>=ES_cutoff) %>% pull(otu.names)])
    df_row_name_label = structure(gsub("D_\\d_", "", rownames(df_row)), names = rownames(df_row))
    ## this is used to create the legend
    LR_sig_range <- df_row$LR %>% range
    
    ## data matrix for the heatmap
    mat_dat <- ft_t_rel_ab_log10 %>% select(Taxa, starts_with("X")) %>% column_to_rownames(var = "Taxa") %>% as.matrix %>% .[rownames(df_row), rownames(df_col)]
    ### annotation components
    mat_col_fun = colorRamp2(c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), c("#253494","#2c7fb8","#41b6c4", "#7fcdbb", "#c7e9b4", "#ffffcc"))
    genus_text_col = ifelse(df_row>0, "#fc8d62", "#66c2a5")
    column_ha = HeatmapAnnotation(df = df_col %>% select(Urban, Age_grp2), col = list(Urban = Urban_color, Age_grp2 = c("Age_0-3" = "#f1eef6", "Age_3-8" = "#bdc9e1","Age_8-18" = "#74a9cf", "Adults" = "#0570b0")), simple_anno_size = unit(3, "mm"), annotation_label = c("Exposure", "Age"), annotation_name_gp = gpar(fontsize = 7), annotation_name_side = "left", annotation_legend_param = list(Urban = list(labels_gp = gpar(fontsize = 7), title_gp = gpar(fontsize = 8, fontface = "bold")), Age_grp2 = list(labels_gp = gpar(fontsize = 7), title_gp = gpar(fontsize = 8, fontface = "bold"))))
    row_ha = rowAnnotation(df = df_row, col = list(LR = colorRamp2(c(-max(abs(LR_sig_range)), 0, max(abs(LR_sig_range))), c("#1b9e77", "white", "#d95f02"))), gp = gpar(col = "black", lwd = 0.3), annotation_name_gp = gpar(fontsize = 7), annotation_name_side = "top", annotation_legend_param = list(LR = list(labels_gp = gpar(fontsize = 7), title_gp = gpar(fontsize = 8, fontface = "bold", angle = 0))))
    
    p <- Heatmap(mat_dat, col = mat_col_fun, name = "Relative Abundance", 
                 heatmap_legend_param = list(at = c(-4, -3.3, -2.6, -1.9, -1.2, -0.5), labels = c("0.01%", "0.05%", "0.25%", "1.25%", "6.25%", "31.25%"), labels_gp = gpar(fontsize = 7), title_gp = gpar(fontsize = 8, fontface = "bold")),
                 
                 cluster_columns = F, show_column_names = F,
                 column_order = rownames(df_col),
                 
                 row_names_gp = gpar(col = genus_text_col, fontsize = 7), row_labels = df_row_name_label[rownames(df_row)],
                 row_dend_width = unit(15, "mm"),
                 row_names_max_width = max_text_width(df_row_name_label, gp = gpar(fontsize = 7)),
                 top_annotation = column_ha,
                 right_annotation = row_ha,
                 height = unit(8, "mm")*dim(mat_dat)[1],
                 width = unit(0.8, "mm")*dim(mat_dat)[2], 
    )
    p1 <- draw(p, heatmap_legend_side = "left", annotation_legend_side = "left")
    pdf(file = paste0("output/figures/Ancom_heatmap_sig_log_rel_ab_", BB, "_", YY, ".pdf"), width = ComplexHeatmap:::width(p1) %>% as.numeric()/25.4, height = ComplexHeatmap:::height(p1) %>% as.numeric()/25.4, useDingbats = F)
    draw(p1)
    dev.off()
    
    p_genus_order = rownames(df_row)[row_order(draw(p))]
    p1_dat <- data.frame(Taxa = p_genus_order, stringsAsFactors = F) %>% separate(Taxa, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "\\.D_\\d_", remove = F) %>% mutate(Domain = gsub(x = Domain, pattern = "D_0_", replacement = ""))
    p1_dat$LR <- lr_b_grp[p1_dat$Taxa]
    p1_dat$Flag <- p1_dat$LR>0
    p1_dat$No = rownames(p1_dat)
    p1_dat$BodySite = BB
    p1_dat$Year = YY
    write.table(p1_dat, file = paste0("output/figures/Ancom_sig_log_rel_ab_", BB, "_", YY, "_genera.txt"), sep = "\t", row.names = F, quote = F)
}
```

2. Lefse style tree plot
```{r all_taxa}
for (ss in seq_along(fts)){
    # ss = 3
    # prepare datasets
    ft_t <- fts[[ss]]
    Mf <- fts_mt[[ss]][["Mf"]] %>% filter(SampleID %in% colnames(ft_t))
    AR_g <- fts_mt[[ss]]$AR
    AR_f <- fts_mt[[ss]]$AR_f
    AR_o <- fts_mt[[ss]]$AR_o
    AR_c <- fts_mt[[ss]]$AR_c
    AR_p <- fts_mt[[ss]]$AR_p
    lr_b_grp = lrs_b_grp[[ss]]
    YY = Mf$Year[1]
    BB = Mf$Body_Site[1]
    
    # defining taxa that changed and unchanged based on sig_cutoff
    ind = paste(YY, BB, sep = "-")
    ES_g_cutoff <- seq(0.6, 0.9, 0.1)[sig_cutoff[ind]] * nrow(AR_g)
    
    # tree
    AR <- data.table::rbindlist(list(AR_g, AR_f, AR_o, AR_c, AR_p), idcol = "taxa.glom") %>% mutate(otu.names = as.character(otu.names), V1 = gsub(pattern = "\\.D", replacement = "/D", otu.names)) %>% separate(V1, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "/")
    AR_tre <- merge(All_genus_tbl %>% select(otu.names, Genus_index), AR, by = "otu.names", all.y = T) %>% filter(Domain!="Unassigned", Phylum!="D_1_") %>% mutate(LR = lr_b_grp[otu.names] %>% as.character(), Sig = ifelse(taxa.glom==1, ifelse(W_stat>=ES_g_cutoff, "Yes", "No"), ifelse(detected_0.8, "Yes", "No")), Sig = ifelse(Sig=="Yes", ifelse(LR>0, "Medium", "Low"), Sig))
    
    Tre <- AR_tre %>% mutate(Root = "Life") %>% unite(pathString, Root, Domain:Family, Genus_index, sep = "/", na.rm = T) %>% as.Node(.$pathString) %>% ToNewick()
    ## convert to treeio tree object
    Tre_td <- as.treedata(ape::read.tree(text = Tre))
    
    ## convert to tibble
    Tre_tb <- as_tibble(Tre_td)
    Tre_tb$branch.length[!is.na(Tre_tb$branch.length)] = 10 # remake the branch length
    Tre_tb$branch.length[grepl(pattern = "^G", Tre_tb$label)]=5
    
    Tre_tb$Ancestors <- NA
    for (i in seq(nrow(Tre_tb))){
        AN = c(ancestor(Tre_tb, Tre_tb$node[i])$label, Tre_tb$label[i]) %>% paste(collapse = "/")
        Tre_tb$Ancestors[i] <- AN
    }
    Tre_tb2 <- Tre_tb %>% separate(Ancestors, into = c("V1", "Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "/") %>% select(-V1)
    Tre_tb3 <- Tre_tb2 %>% left_join(., AR_tre, by = c("Domain", "Phylum", "Class", "Order", "Family", "Genus" = "Genus_index")) %>% select(parent, node, branch.length, everything())
    class(Tre_tb3) <- c("tbl_tree", class(Tre_tb3))
    
    ### node to label (1) all abundant phylum
    ft_t_rel_ab_p <- ft_t_rel_ab %>% filter(taxa.glom=="Phylum") %>% gather(key = "SampleID", value = "Rel_ab", starts_with("X"))
    # ggplot(ft_t_rel_ab_p, aes(x = Phylum, y = Rel_ab)) +
    #     geom_boxplot() +
    #     stat_summary(geom = "point", fun = "median", color = "red") +
    #     scale_y_continuous(breaks = seq(0, 0.1, 0.005)) +
    #     coord_flip(ylim = c(0, 0.1))
    Phylum_abundant <- ft_t_rel_ab_p %>% group_by(Phylum) %>% summarise(Med = median(Rel_ab)) %>% filter(Med>0.001) %>% pull(Phylum)
    ### (2) family and phylum of significantly changed genus
    node_sig <- Tre_tb3 %>% filter(!Sig=="No") %>% pull(node)
    node_nosig <- Tre_tb3 %>% filter(Sig=="No"|is.na(Sig)) %>% pull(node)
    node_genus_sig <- Tre_tb3 %>% filter(taxa.glom==1, Sig!="No") %>% pull(node)
    
    node_family_g_sig <- sapply(node_genus_sig, function(x){parent(Tre_tb3, x)$node})
    node_phylum_g_sig <- sapply(node_genus_sig, function(x){ancestor(Tre_tb3, x) %>% filter(grepl("D_1_", label, fixed = T)) %>% pull(node)})
    ### (3) clade with no significantly changed child
    node_clade_nosig <- node_nosig[!sapply(node_nosig, function(x){any(offspring(Tre_tb3, x)$node %in% node_sig)})]
    node_clade_nosig_top <- node_clade_nosig[!sapply(node_clade_nosig, function(x){any(ancestor(Tre_tb3, x)$node %in% node_clade_nosig)})]
    node_clade_nosig_top_p <- sapply(node_clade_nosig_top, function(x){parent(Tre_tb3, x)})
    ## (4) significantly changed node but without any significantly changed offspring node
    node_sig_no_offspring <- node_sig[!sapply(node_sig, function(x){any(offspring(Tre_tb3, x)$node %in% node_sig)})]
    
    ### node to not label (1) intermediate of a series of taxa
    node_sig_int <- node_sig[sapply(node_sig, function(x){any(ancestor(Tre_tb3, x)$node %in% node_sig) & (Tre_tb3$taxa.glom[Tre_tb3$node == x] %in% c(3, 4))})]
    
    node_to_label = c(Tre_tb3$node[Tre_tb3$label %in% c(Phylum_abundant, "D_0_Archaea", "D_0_Bacteria")], node_family_g_sig, node_phylum_g_sig, node_clade_nosig_top_p, node_sig_no_offspring) %>% unique %>% .[!. %in% node_sig_int]
    
    
    Tre_tb4 <- Tre_tb3 %>% mutate(label2 = ifelse(node %in% node_to_label, gsub("D_\\d_", "", label), NA_character_)) %>% arrange(node)
    
    Tre_td_new <- Tre_tb4 %>% as.treedata()
    p <- ggtree(Tre_td_new, layout='circular', size = 0.3) + 
        geom_point2(aes(subset = Sig=="No"), size = 0.5, fill = "yellow", color = "black", shape = 21) +
        geom_point2(aes(subset = Sig!="No", fill = Sig, size = abs(as.numeric(LR))), color = "black", shape = 21) +
        geom_text2(aes(subset = !isTip, x = branch, label = label2, angle = ifelse((angle < 90 | angle >270), angle, angle + 180)), nudge_y = -0.5, size = 4/.pt) +
        geom_tiplab2(aes(subset = Sig!="No", label = Genus.y), size = 5/.pt, offset = 2) +
        scale_fill_manual(values = Urban_color) +
        scale_size_continuous(breaks = c(1, 2, 3), labels = c(2, 4, 8), range = c(1.5, 6)) +
        labs(fill = "Exposure", size = "Fold change")
    
    Tre_tb_new_sub <- Tre_tb4 %>% filter(taxa.glom!=1, Sig != "No")
    for (i in seq(nrow(Tre_tb_new_sub))){
        NN = Tre_tb_new_sub[i, ]$node
        SS = Tre_tb_new_sub[i, ]$Sig
        p = p + geom_hilight(node = NN, fill = Urban_color[SS], alpha = 0.3)
    }

    ggsave(plot = p, paste0("output/figures/Ancom_tree_", BB, "_", YY, "_v1.pdf"), width = 11, useDingbats = F)
    
    for (NN2 in c(node_clade_nosig_top, node_sig_no_offspring)){
        p = collapse(p, NN2)
    }
    ggsave(plot = p, paste0("output/figures/Ancom_tree_", BB, "_", YY, "_v2.pdf"), width = 11, useDingbats = F)
}
```

```{r Sig_taxa}
for (ss in seq_along(fts)){
    # ss = 3
    # prepare datasets
    ft_t <- fts[[ss]]
    Mf <- fts_mt[[ss]][["Mf"]] %>% filter(SampleID %in% colnames(ft_t))
    AR_g <- fts_mt[[ss]]$AR
    AR_f <- fts_mt[[ss]]$AR_f
    AR_o <- fts_mt[[ss]]$AR_o
    AR_c <- fts_mt[[ss]]$AR_c
    AR_p <- fts_mt[[ss]]$AR_p
    lr_b_grp = lrs_b_grp[[ss]]
    YY = Mf$Year[1]
    BB = Mf$Body_Site[1]
    
    # defining genus that changed and unchanged based on sig_cutoff
    ind = paste(YY, BB, sep = "-")
    ES_g_cutoff <- seq(0.6, 0.9, 0.1)[sig_cutoff[ind]] * nrow(AR_g)
    
    ## prepare the metadata
    AR <- data.table::rbindlist(list(AR_g, AR_f, AR_o, AR_c, AR_p), idcol = "taxa.glom") %>% mutate(otu.names = as.character(otu.names), V1 = gsub(pattern = "\\.D", replacement = "/D", otu.names)) %>% separate(V1, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "/")
    
    AR_tre <- merge(All_genus_tbl %>% select(otu.names, Genus_index), AR, by = "otu.names", all.y = T) %>% filter(Domain!="Unassigned", Phylum!="D_1_") %>% mutate(LR = lr_b_grp[otu.names] %>% as.character(), Sig = ifelse(taxa.glom==1, ifelse(W_stat>=ES_g_cutoff, "Yes", "No"), ifelse(detected_0.8, "Yes", "No")), Sig = ifelse(Sig=="Yes", ifelse(LR>0, "Medium", "Low"), Sig)) %>% filter(Sig!="No")
    
    # tree
    Tre <- AR_tre %>% mutate(Root = "Life") %>% unite(pathString, Root, Domain:Family, Genus_index, sep = "/", na.rm = T) %>% as.Node(.$pathString) %>% ToNewick()

    ## convert to treeio tree object
    Tre_td <- as.treedata(ape::read.tree(text = Tre))
    ## convert to tibble
    Tre_tb <- as_tibble(Tre_td)
    Tre_tb$branch.length[!is.na(Tre_tb$branch.length)] = 10 # remake the branch length
    Tre_tb$branch.length[grepl(pattern = "^G", Tre_tb$label)]=5
    
    Tre_tb$Ancestors <- NA
    
    for (i in seq(nrow(Tre_tb))){
        AN = c(ancestor(Tre_tb, Tre_tb$node[i])$label, Tre_tb$label[i]) %>% paste(collapse = "/")
        Tre_tb$Ancestors[i] <- AN
    }
    Tre_tb2 <- Tre_tb %>% separate(Ancestors, into = c("V1", "Domain", "Phylum", "Class", "Order", "Family", "Genus"), sep = "/", remove = F) %>% select(-V1)
    

    Tre_tb3 <- Tre_tb2 %>% left_join(., AR_tre, by = c("Domain", "Phylum", "Class", "Order", "Family", "Genus" = "Genus_index")) %>% select(parent, node, branch.length, everything())
    class(Tre_tb3) <- c("tbl_tree", class(Tre_tb3))
    
    ### node to label (1) family of significantly changed genus
    node_sig <- Tre_tb3 %>% filter(!Sig=="No") %>% pull(node)
    node_nosig <- Tre_tb3 %>% filter(Sig=="No"|is.na(Sig)) %>% pull(node)
    node_genus_sig <- Tre_tb3 %>% filter(taxa.glom==1, Sig!="No") %>% pull(node)
    
    node_family_g_sig <- sapply(node_genus_sig, function(x){parent(Tre_tb3, x)$node})
    
    ## (3) significantly changed node but without any significantly changed offspring node
    node_sig_no_offspring <- node_sig[!sapply(node_sig, function(x){any(offspring(Tre_tb3, x)$node %in% node_sig)})]
    
    ### node to not label (1) intermediate of a series of taxa
    node_sig_int <- node_sig[sapply(node_sig, function(x){any(ancestor(Tre_tb3, x)$node %in% node_sig) & (Tre_tb3$taxa.glom[Tre_tb3$node == x] %in% c(3, 4))})]
    
    node_to_label = c(Tre_tb3$node[Tre_tb3$label %in% c("D_0_Archaea", "D_0_Bacteria")], node_family_g_sig, node_sig_no_offspring) %>% unique %>% .[!. %in% node_sig_int]
    
    Tre_tb4 <- Tre_tb3 %>% mutate(label2 = ifelse(node %in% node_to_label, gsub("D_\\d_", "", label), NA_character_)) %>% arrange(node)
    
    Tre_td_new <- Tre_tb4 %>% as.treedata()
    p <- ggtree(Tre_td_new, layout='circular', size = 0.3) + 
        geom_point2(aes(subset = is.na(Sig)), size = 0.5, fill = "yellow", color = "black", shape = 21) +
        geom_point2(aes(subset = !is.na(Sig), fill = Sig, size = abs(as.numeric(LR))), color = "black", shape = 21) +
        geom_text2(aes(subset = !isTip, x = branch, label = label2, angle = ifelse((angle < 90 | angle >270), angle, angle + 180)), nudge_y = -0.2, size = 5/.pt) +
        geom_text2(aes(subset = grepl("D_1_", label, fixed = T), x = branch, label = gsub("D_\\d_", "", Phylum), angle = ifelse((angle < 90 | angle >270), angle, angle + 180)), nudge_y = -0.3, size = 5/.pt) +
        geom_tiplab2(aes(subset = Sig!="No", label = Genus.y), size = 5/.pt, offset = 2) +
        scale_fill_manual(values = Urban_color) +
        scale_size_continuous(breaks = c(1, 2, 3), labels = c(2, 4, 8), range = c(1.5, 6)) +
        labs(fill = "Exposure", size = "Fold change")
    p
    
    ggsave(plot = p, paste0("output/figures/Ancom_tree_", BB, "_", YY, "_v3.pdf"), width = 11, useDingbats = F)
    
    for (NN2 in node_sig_no_offspring){
        p = collapse(p, NN2)
    }
    ggsave(plot = p, paste0("output/figures/Ancom_tree_", BB, "_", YY, "_v4.pdf"), width = 11, useDingbats = F)
}
```



### Songbirds
Preparation of biom and metadata for Songbirds analysis in bash
```{r, eval=FALSE}
ft_g <- ft_glom$Genus
genus_tbl <- data.frame(Genus = ft_g$Taxa, Index = paste0("Silva_Genus_", seq(nrow(ft_g))))
ft_g2 <- ft_g[, -c(1:7)] %>% as.matrix()
rownames(ft_g2) <- genus_tbl$Index

ob_mt_g <- genus_tbl %>% separate(Genus, into = c("kingdom", "phylum", "class", "order", "family", "genus"), sep = "\\|") %>% column_to_rownames(var = "Index") %>% as.matrix()



for (YY in Yrs){
    #YY = "2015"
    for (BB in BodySites){
        # BB = "Feces"
        Mf <- Mapping_work2 %>% filter(Year==YY, Body_Site==BB, SampleGroup=="Villagers")
        Mf_lst <- Mf %>% split(Mf$Age_grp2)
        
        biom_lst <- lapply(Mf_lst, function(x){create_biom(x, ft_g2)})
        
        # write biom files
        lapply(seq_along(biom_lst), function(x){write_biom(biom_lst[[x]][[1]], paste0("misc/Songbirds/data/ft_", names(biom_lst)[x], "_", BB, "_", YY, ".biom"))})
        # write metadata
        lapply(seq_along(biom_lst), function(x){write.table(biom_lst[[x]][[2]], file = paste0("misc/Songbirds/data/mt_", names(biom_lst)[x], "_", BB, "_", YY, ".txt"), sep = "\t", quote = F, row.names = F)})
    }
}
```
```{shell}
cd Songbirds
conda activate songbird_env

for yy in 2015 2016
do
    for bb in Feces Mouth Nose Right_Arm Right_Hand
    do
        for f in Age_0-3 Age_3-8 Age_8-18 Adults; do songbird multinomial --input-biom data/ft_${f}_${bb}_${yy}.biom --metadata-file data/mt_${f}_${bb}_${yy}.txt --formula "C(urban, Treatment('Low'))+gender" --epochs 10000 --differential-prior 0.5 --summary-interval 10 --summary-dir ${yy}_${bb}_${f}_results --random-seed 42; done &> ${bb}_${yy}.log &
    done
done

```


### Maaslin2


<!-- ```{r} -->
<!-- test_sum <- test_ancom_ad_15$W.taxa %>% filter(detected_0.9) %>% select(otu.names) %>% mutate(Sig1 = "Adults_15") %>% merge(., test_ancom_ch_15$W.taxa %>% filter(detected_0.9) %>% select(otu.names) %>% mutate(Sig2 = "Children_15"), by = "otu.names", all = T) %>% merge(., test_ancom_ad_16$W.taxa %>% filter(detected_0.9) %>% select(otu.names) %>% mutate(Sig3 = "Adults_16"), by = "otu.names", all = T) %>% merge(., test_ancom_ch_16$W.taxa %>% filter(detected_0.9) %>% select(otu.names) %>% mutate(Sig4 = "Children_16"), by = "otu.names", all = T) -->

<!-- ft_picked <- ft_genus %>% filter(Taxa_id %in% test_sum$otu.names) -->
<!-- ft_picked_long <- ft_picked %>% select(Taxa, Taxa_id, everything()) %>% gather(key = "SampleID", value = "Count", -c(1, 2)) %>% merge(., Work_Map, by = "SampleID") %>% mutate(plot_grp = paste0(Year, Age_grp1)) -->

<!-- ggplot(ft_picked_long, aes(x = plot_grp, y = Count, color = Urban)) + -->
<!--     geom_boxplot() + -->
<!--     facet_wrap(~Taxa_id, scales = "free_y") -->
<!-- ``` -->

