---
title: "Alpha Diversity"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE, warning=FALSE}
library(knitr)
options(max.print="75")
opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, prompt=FALSE, tidy = TRUE, comment = NA, message = FALSE)
opts_knit$set(width=75)

# Loading library here
library(tidyverse)
library(ggpubr)
library(MASS)
library(segmented)
library(agricolae)
library(car)
library(effectsize)
library(rstatix)
library(patchwork)

# set path
source("../code/utils.R")

```
## Initial setup
**Setup color index**
```{r}
# Exposure level
Expo_color <- c("#66c2a5", "#fc8d62", "#8da0cb")
names(Expo_color) <- c("Low", "Medium", "High")
print(Expo_color)
#plot(1:3, 1:3, col = Expo_color, pch = 16, cex = 4)
# Ethnicity
Ethnicity_color <- gg_color_hue(3)
names(Ethnicity_color) <- c("SANEMA", "YEKWANA", "Visitors")
print(Ethnicity_color)

Sanemas <- c("Chajuranha", "Mosenahanha", "Kuyuwininha", "Shianana-Jiyakwanha", "Washudihanha", "Sudukuma")
Yekwana <- c("Kanarakuni", "Fiyakwanha")
```

**Import the data**
```{r}
# Alpha diversities
alpha = readRDS("../data/alpha_imported_all.RDS")
```
```{r class.source='fold-show'}
# Mapping
## This mapping files include all villagers and baseline of visitors of all body sites
mt_ms = readRDS("../data/Mapping_MS_Expo.Rds")
Mapping_work <- mt_ms %>% mutate(Expo = ifelse(Ethnicity=="SANEMA", "Low", ifelse(Village=="Fiyakwanha", "Low", ifelse(SampleGroup=="Visitors", "High", "Medium"))) %>% factor(levels = c("Low", "Medium", "High")), Age_num = as.numeric(Age), Age_num = ifelse(is.na(Age_num), as.numeric(gsub(pattern = "_months", replacement = "", Age, ignore.case = T))/12, Age_num), Age_grp1 = ifelse(Age_num>=18, "Adults", "Children"))
Mapping_work$Age_num[Mapping_work$Age=="4_Days"] <- 0
Mapping_work$Age_num[Mapping_work$Age=="1_day"] <- 0
```

**Set up the Analysis**
```{r}
dat_alpha <- merge(alpha, Mapping_work, by = 1) %>% mutate(Age_grp2 = case_when(Age_num<=3 ~ "Age_0-3", Age_num<=8 ~ "Age_3-8", Age_num<18 ~ "Age_8-18", T ~ "Adults"))
```

## Effects of exposure level
```{r}
Indices <- colnames(alpha[c(2:5)])
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")
```
### Initial test on feces (no need to run)
**We first check the faith_pd which described the overall complexity of the microbial community.**
```{r, eval=FALSE}
BB = "Feces"
AA = "faith_pd"
# Year 2015
YY = "2015"
Work_dat <- dat_alpha %>% dplyr::filter(Body_Site==BB, Year==YY, SampleGroup=="Villagers")

# Obtain the full linear model
fit.full <- lm(get(AA) ~ Age_num + Expo + Gender + Age_num:Expo + Age_num:Gender + Expo:Gender, data = Work_dat)

# Obtain the null model with no predictors, this is necessary for forward selection use stepAIC
fit.null <- lm(get(AA) ~ 1, data = Work_dat)

# Forward selection, scope parameters provides the full model
regforward.out <- stepAIC(fit.null, direction = "forward", trace = 4, scope = ~ Age_num + Expo + Gender + Age_num:Expo + Age_num:Gender + Expo:Gender)

anova(regforward.out)
```
The best model from 2015 includes Expo, Age, Expo:Age interaction and Gender (in decreasing importance).

```{r, eval=FALSE}
# Year 2016
YY = "2016"
Work_dat <- Work %>% dplyr::filter(Body_Site==BB, Year==YY, SampleGroup=="Villagers")

# Obtain the full linear model
fit.full <- lm(get(AA) ~ Age_num + Expo + Gender + Age_num:Expo + Age_num:Gender + Expo:Gender, data = Work_dat)

# Obtain the null model with no predictors, this is necessary for forward selection use stepAIC
fit.null <- lm(get(AA) ~ 1, data = Work_dat)

# Forward selection, scope parameters provides the full model
regforward.out <- stepAIC(fit.null, direction = "forward", trace = 4, scope = ~ Age_num + Expo + Gender + Age_num:Expo + Age_num:Gender + Expo:Gender)

anova(regforward.out)
```
The best model from 2016 includes Expo, Age, Expo:Age interaction (in decreasing importance), but Gender was not there.


So we will fitting the data with Expo, Age, Expo:Age interaction, and Gender for both years to keep unity.
```{r, eval=FALSE}
# Final model
lm_models <- list()
for (YY in Yrs){
  Work_dat <- Work %>% dplyr::filter(Body_Site==BB, Year==YY, SampleGroup=="Villagers")
  fit.final <- lm(as.formula(paste0(AA, " ~ Age_num + Expo + Age_num:Expo + Gender")), data = Work_dat)
  print(paste0("--------Below is ANOVA table for year ", YY))
  anova(fit.final) %>% print()
  
  print(paste0("--------Below is coefficient for year ", YY))
  summary(fit.final) %>% print()
  lm_models[[YY]] <- fit.final
}
```

Next we perform the broken stick regression based on the selected final model
```{r, eval=FALSE}
for (YY in Yrs){
  #YY = "2016"
  fit.seg <- segmented(lm_models[[YY]], seg.Z = ~ Age_num, psi = c(3))
  summary(fit.seg)
  plot(fit.seg$fitted.values, fit.seg$residuals)
  plot(fit.seg$model$faith_pd, fit.seg$fitted.values)
  plot(fit.seg$model$Age_num, hatvalues(fit.seg))
  plot(fit.seg$model$Age_num, cooks.distance(fit.seg))
  plot(hatvalues(fit.seg), fit.seg$residuals)
}
```

It seems from the residual plots that the point with faith_pd smaller than 10 is dragging the curve down and causing non-random residual. 

```{r, eval=FALSE}
Work %>% dplyr::filter(Body_Site==BB, faith_pd<12, SampleGroup=="Villagers")
```
It does appears that many baby are outliers, so re-do the analysis without baby.

### Modeling without babies
#### Overview
```{r}
for (bb in BodySites){
    for (aa in Indices){
        Work_dat <- dat_alpha %>% dplyr::filter(Body_Site==bb, SampleGroup=="Villagers", Age_num>=1)
        p <- ggplot(Work_dat, aes(x = Age_num, y = get(aa), color = Expo, shape = Gender)) +
            geom_point(size = 1.5) +
            scale_shape_manual(values = c(16, 21)) +
            scale_color_manual(breaks = names(Expo_color), values = Expo_color) +
            facet_grid(.~Year) +
            theme_jw() + theme(aspect.ratio = 0.8, panel.background = element_rect(color = "black")) +
            labs(x = "Age (Years)", y = aa, title = bb)
        print(p)
    }
}
```
#### Statisticals
##### Children
Building full multiple linear regression models with interaction terms for children
```{r}
alpha_youngchild_lm_models = list()
lm.stats.tbl = list()
for (bb in BodySites){
    for (aa in Indices){
        for (yy in Yrs){
            Work_dat <- dat_alpha %>% dplyr::filter(Body_Site==bb, Year==yy, SampleGroup=="Villagers", Age_num>=1, Age_num<18)
            lm.model <- lm(paste0(aa, " ~ Age_num + Expo + Gender + Age_num:Expo + Age_num:Gender + Expo:Gender"), data = Work_dat, contrasts = list(Expo = "contr.sum", Gender = "contr.sum"))
            lm.stats = Anova(lm.model, type = 3) %>% # Type 3 ss
                merge(., effectsize::eta_squared(., ci = 0.95), by.x = 0, by.y = 1, all = T) %>% # effect size estimates
                rename(Factors = Row.names) %>% mutate(Year = yy, Body_Site = bb, Indice = aa)
            alpha_youngchild_lm_models[[bb]][[aa]][[yy]] = lm.model
            lm.stats.tbl[[paste(bb, aa, yy, sep = "-")]] = lm.stats
        }
    }
}
lm.stats.tbls = do.call("rbind", lm.stats.tbl) %>% rename(p_value = `Pr(>F)`) %>% mutate(q_value = p.adjust(p_value, method = "fdr"))
write.table(lm.stats.tbls, file = "../data/alpha_lm_stats_children.txt", sep = "\t", row.names = F, quote = F)

lm.stats.tbls.2 = lm.stats.tbls %>% dplyr::filter(!is.na(Eta2_partial)) %>% mutate(Factors = factor(Factors, levels = c("Age_num", "Gender", "Expo", "Age_num:Expo", "Age_num:Gender", "Expo:Gender"))) %>% arrange(Body_Site, Indice, Year, Factors)
    
ggplot(lm.stats.tbls.2, aes(x = Factors, y = q_value, color = Body_Site)) +
    geom_point() +
    geom_path(aes(group = Year, linetype = Year)) +
    facet_grid(Body_Site~Indice) +
    scale_y_continuous(breaks = c(0, 0.05, 0.1, 0.2, 1.0)) +
    theme_jw(base_size = 12) + theme(aspect.ratio = 0.4, axis.text.x = element_text(angle = 90, hjust = 1), panel.grid.minor = element_blank())

ggplot(lm.stats.tbls.2, aes(x = Factors, y = Eta2_partial, color = Body_Site)) +
    geom_point() +
    geom_path(aes(group = Year, linetype = Year)) +
    facet_grid(Body_Site~Indice, scales = "free_y") +
    #scale_y_continuous(breaks = c(0, 0.05, 0.1, 0.2, 1.0)) +
    theme_jw() + theme(aspect.ratio = 0.4, axis.text.x = element_text(angle = 90, hjust = 1))
```
No strong interactions obsereved among children

##### all ages
```{r}
alpha_full_lm_models = list()
lm.stats.tbl = list()
for (bb in BodySites){
    for (aa in Indices){
        for (yy in Yrs){
            Work_dat <- Work %>% dplyr::filter(Body_Site==bb, Year==yy, SampleGroup=="Villagers", Age_num>=1)
            lm.model <- lm(paste0(aa, " ~ Age_num + Expo + Gender"), data = Work_dat, contrasts = list(Expo = "contr.sum", Gender = "contr.sum"))
            lm.stats = Anova(lm.model, type = 3) %>% # Type 3 ss
                merge(., effectsize::eta_squared(., ci = 0.95), by.x = 0, by.y = 1, all = T) %>% # effect size estimates
                rename(Factors = Row.names) %>% mutate(Year = yy, Body_Site = bb, Indice = aa)
            alpha_full_lm_models[[bb]][[aa]][[yy]] = lm.model
            lm.stats.tbl[[paste(bb, aa, yy, sep = "-")]] = lm.stats
        }
    }
}
lm.stats.tbl = do.call("rbind", lm.stats.tbl)
write.table(lm.stats.tbl, file = "../data/alpha_lm_stats.txt", sep = "\t", row.names = F, quote = F)

lm.stats.tbl.2 = lm.stats.tbl %>% dplyr::filter(!is.na(Eta2_partial)) %>% rename(p_value = `Pr(>F)`) %>% mutate(Factors = factor(Factors, levels = c("Age_num", "Gender", "Expo"))) %>% arrange(Body_Site, Indice, Year, Factors)
    
p1 = ggplot(lm.stats.tbl.2 %>% dplyr::filter(!Indice %in% c("observed_otus", "shannon")), aes(x = Factors, y = p_value, color = Body_Site)) +
    geom_point() +
    geom_path(aes(group = Year, linetype = Year)) +
    facet_grid(Body_Site~Indice) +
    scale_y_continuous(breaks = c(0, 0.05, 0.1, 0.2, 1.0)) +
    theme_jw() + theme(aspect.ratio = 0.6, axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(color = F) +
    ggtitle("p_value")

p2 = ggplot(lm.stats.tbl.2 %>% dplyr::filter(!Indice %in% c("observed_otus", "shannon")), aes(x = Factors, y = Eta2_partial, color = Body_Site)) +
    geom_point() +
    geom_path(aes(group = Year, linetype = Year)) +
    facet_grid(Body_Site~Indice, scales = "free_y") +
    theme_jw() + theme(aspect.ratio = 0.6, axis.text.x = element_text(angle = 90, hjust = 1)) +
    guides(color = F) +
    ggtitle("Eta2")

p1 + p2 + plot_layout(guides = "collect")
ggsave("fig_tmp/alpha_stats.pdf", width = 9)
```

For fecal microbiome, exposure, gender, and age are all significant in 2015, with exposure remain significant for 2016 for faith_pd, but not for pielou_e. The effect size of exposure are high, with 2016 has a smaller effect size. The difference seen in faith pd may be due to small number of less abundant species.

For mouth and nasal microbiome, the pattern are quite similar, both age and exposure have strong effects to faith_pd, but not for pielou'e. Again, 2016 seems to have less effect of exposure.

For the two skin sites, exposure has a strong effect for both sites in both years to faith_pd. Not pielou's e.

#### Models
```{r}
models = list()
model_tbls = data.frame(BodySite = as.character(),
                        Index = as.character(),
                        Year = as.character(),
                        Model = as.character(),
                        Model.type = as.character())

for (bb in BodySites){
    for (aa in Indices){
        for (yy in Yrs){
            Work_dat <- Work %>% dplyr::filter(Body_Site==bb, Year==yy, SampleGroup=="Villagers", Age_num>=1)
            # Obtain the full linear model
            fit.full <- lm(paste0(aa, " ~ Age_num + Expo + Gender"), data = Work_dat)
            # Obtain the null model with no predictors, this is necessary for forward selection use stepAIC
            fit.null <- lm(paste0(aa, " ~ 1"), data = Work_dat)
            
            # Forward selection, scope parameters provides the full model
            regforward.out <- stepAIC(fit.null, direction = "forward", trace = 4, scope = ~ Age_num + Expo + Gender)
            
            lm.model <- lm(regforward.out$terms, data = Work_dat, contrasts = list(Expo = "contr.sum", Gender = "contr.sum"))
            lm.model.anova = car::Anova(lm.model, type = 3)
            
            if ("Age_num" %in% rownames(lm.model.anova)){
                fit.seg <- segmented(regforward.out, seg.Z = ~ Age_num, psi = 9)
                model_comp = anova(fit.seg, lm.model)
                if (!is.na(model_comp[2, 6]) & model_comp[2, 6]<=0.05){
                    fit.final = fit.seg
                    summary(fit.seg) %>% print()
                    davies.test(lm.model, seg.Z = ~ Age_num, k = 10) %>% print()
                    confint.segmented(fit.seg) %>% print()
                    slope(fit.seg) %>% print()
                } else {
                    fit.final = lm.model
                }
            } else {
                fit.final = lm.model
            }
            plot(fit.final$fitted.values, fit.final$residuals)
            plot((fit.final$model)[, 1], fit.final$fitted.values)
            plot(hatvalues(fit.final), fit.final$residuals)
            qqnorm(fit.final$residuals)
            ff = (fit.final$terms %>% as.character())[3]
            models[[bb]][[aa]][[yy]] = fit.final
            if ("segmented" %in% class(fit.final)){
                model_tbls = rbind(model_tbls, cbind(bb, aa, yy, ff, "segmented"))
            } else {
                model_tbls = rbind(model_tbls, cbind(bb, aa, yy, ff, "linear"))
            }
        }
    }
}
colnames(model_tbls) = c("Body_Site", "Indice", "Year", "Model", "Model.type")
```
#### Graphs
```{r}
Indices <- colnames(alpha[c(2:5)])
Yrs <- c("2015", "2016")
BodySites <- c("Feces", "Nose", "Mouth", "Right_Arm", "Right_Hand")

dat_new = complete(cbind(Age_num = seq(1, 60, 1), 
                         Expo = c("Low", "Medium"), 
                         Gender = c("F", "M")) %>% 
                       as.data.frame(), Age_num, Expo, Gender) %>% 
    mutate(Age_num = as.numeric(Age_num))

dat_pred = list()
for (bb in BodySites){
    for (aa in Indices){
        for(yy in Yrs){
            dat_pred[[bb]][[aa]][[yy]] = cbind(dat_new, predict(models[[bb]][[aa]][[yy]], newdata = dat_new, interval = "confidence"))
            
        }
    }
}
for (yy in Yrs){
    for (aa in Indices){
        #aa = "faith_pd"
        p_list = list()
        for (bb in BodySites){
            #bb = "Feces"
            dat_ = Work %>% dplyr::filter(Body_Site==bb, Year==yy, SampleGroup=="Villagers", Age_num>=1)
            dat_pred_ = dat_pred[[bb]][[aa]][[yy]]
            mod = model_tbls %>% dplyr::filter(Body_Site==bb, Year==yy, Indice==aa) %>% pull(Model)
            p = ggplot(dat_, aes_string(x = "Age_num", y = aa, color = "Expo", shape = "Gender")) +
                geom_point(size = 1, alpha = 0.5) +
                geom_line(data = dat_pred_, aes(y = fit, linetype = Gender), size = 0.3) +
                geom_ribbon(data = dat_pred_, aes(y = fit, ymin = lwr, ymax = upr, fill = Expo, alpha = Gender), color = NA, show.legend = F) +
                annotate("label", x = Inf, y = Inf, label = str_wrap(paste0(aa, " ~ ", mod), width = 50), hjust = 1, vjust = 1, size = 8/.pt) +
                scale_color_manual(values = Expo_color) +
                scale_fill_manual(values = Expo_color) +
                scale_shape_manual(values = c(16, 1)) +
                scale_linetype_manual(values = c(1, 2)) +
                scale_alpha_manual(values = c(0.4, 0.2)) +
                theme_jw(base_size = 10) + theme(aspect.ratio = 0.7, panel.background = element_rect(color = "black"), plot.title = element_text(hjust = 0, vjust = 1), plot.title.position = "panel", legend.key.width = unit(1,"cm"), legend.margin = margin(rep(-3, 4))) +
                labs(color = "Samples points", linetype = "Regression lines", shape = "", x = "Age (Year)") +
                guides(color = guide_legend(override.aes = list(alpha = 1, size = 1.5, linetype = c(NA, NA)), order = 1), shape = guide_legend(override.aes = list(size = 1.5, alpha = 1), order = 2), linetype = guide_legend(order = 3)) +
                ggtitle(bb)
            p_list[[bb]] = p
        }
        p_f = p_list[["Feces"]] + p_list[["Mouth"]] + guide_area() + p_list[["Nose"]] + p_list[["Right_Arm"]] + p_list[["Right_Hand"]] + plot_layout(guides = "collect", nrow = 2) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 10, face = "bold"))
        print(p_f)
        set_panel_size(p = p_f, g = patchworkGrob(p_f), file = paste0( "../output/figures/alpha_regression_", aa, "_", yy, ".pdf"), width =  unit(3, "in"), height = unit(2.1, "in"))
    }
}
```

## Effects of year

```{r}
Indices <- colnames(alpha[c(2:5)])
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")
```

```{r}
for (bb in BodySites){
    for (aa in Indices){
        #break
        dat_ = Work %>% dplyr::filter(Body_Site==bb, SampleGroup=="Villagers", Age_num>=1) %>% mutate(Year = as.character(Year), Age_grp2 = factor(Age_grp2, levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")))
        
        dat_.test = dat_ %>% group_by(Expo, Age_grp2) %>% wilcox_test(formula(paste0(aa, "~ Year"))) %>% add_xy_position(x = "Age_grp2", dodge = 0.7) %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "fdr") %>% round(digits = 3), p.adj.signif = ifelse(p.adj<=0.1, p.adj, "ns"))
        
        ggplot(dat_, aes_string(x = "Age_grp2", y = aa, color = "Year")) +
            geom_point(size = 0.75, position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.7)) +
            geom_boxplot(aes(fill = Year), width = 0.5, color = "black", position = position_dodge(width = 0.7), size = 0.2, outlier.alpha = 0) +
            stat_pvalue_manual(dat_.test, label = "p.adj", tip.length = 0, size = 8/.pt) +
            facet_grid(.~Expo) +
            theme_jw(base_size = 8) + theme(aspect.ratio = 0.8, legend.position = "top", legend.direction = "horizontal", panel.background = element_rect(color = "black"), plot.title = element_text(hjust = 0, vjust = 1), plot.title.position = "panel") +
            scale_fill_manual(values = c(NA, NA)) +
            scale_color_manual(values = c("#80b1d3", "#fb8072")) +
            labs(x = "") + ggtitle(bb)
        set_panel_size(p = last_plot(), file = paste0( "fig_tmp/alpha_compare_year_", aa, "_", bb, ".pdf"), width =  unit(3, "in"), height = unit(2.4, "in"))
    }
}
```

## Delta alpha diversity
```{r}
Indices <- colnames(alpha[c(2:5)])
BodySites <- c("Feces", "Mouth", "Nose", "Right_Arm", "Right_Hand")
```

```{r}
p_list = list()
for (bb in BodySites){
    for (aa in Indices){
        #break
        dat_ = Work %>% dplyr::filter(Body_Site==bb, SampleGroup=="Villagers", Age_num>=1) %>% mutate(Year = as.character(Year), Age_grp2 = factor(Age_grp2, levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Expo = as.character(Expo))
        d_alpha = split(dat_, list(dat_$Year, dat_$Age_grp2, dat_$Expo)) %>%
            lapply(., function(x){dist(x[[aa]], method = "manhattan") %>% as.numeric}) %>% do.call("c", .) %>% data.frame(delta_alpha = .) %>% rownames_to_column(var = "Tmp") %>% separate(Tmp, c("Year", "Age_grp2", "Expo"), sep = "\\.") %>% mutate(Expo = gsub("\\d+", "", Expo), Age_grp2 = factor(Age_grp2, levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")))
        d_alpha.test = d_alpha %>% group_by(Year, Age_grp2) %>% wilcox_test(delta_alpha ~ Expo) %>% add_xy_position(x = "Age_grp2", dodge = 0.7) %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "fdr") %>% round(digits = 3), p.adj.signif = ifelse(p.adj<=0.1, p.adj, "ns"))
        
        p_list[[bb]][[aa]] <- ggplot(d_alpha, aes(x = Age_grp2, y = delta_alpha)) +
            geom_quasirandom(size = 1.5, aes(fill = Expo), dodge.width = 0.7, shape = 21, alpha = 0.5) +
            geom_boxplot(width = 0.5, aes(color = Expo), position = position_dodge(width = 0.7), fill = NA, size = 0.3, outlier.alpha = 0) +
            stat_pvalue_manual(d_alpha.test, label = "p.adj", tip.length = 0, size = 8/.pt) +
            facet_grid(.~Year) +
            scale_fill_manual(values = Expo_color) +
            scale_color_manual(values = c("black", "black")) +
            theme_jw(base_size = 8) + theme(aspect.ratio = 0.8) +
            labs(y = paste0("Absolute value \u0394(", aa, ")")) + ggtitle(bb)
        print(p_list[[bb]][[aa]])
        set_panel_size(p = p_list[[bb]][[aa]], file = paste0("fig_tmp/delta_alpha_expo_", aa, "_", bb, ".pdf"), width = unit(3, "in"), height = unit(2.4, "in"))
    }
}

p_list[["Feces"]][["faith_pd"]]+p_list[["Feces"]][["pielou_e"]] + plot_layout(guides = "collect", ncol = 1)

for (bb in BodySites){
    for (aa in Indices){
        #break
        dat_ = Work %>% dplyr::filter(Body_Site==bb, SampleGroup=="Villagers", Age_num>=1) %>% mutate(Year = as.character(Year), Age_grp2 = factor(Age_grp2, levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")), Expo = as.character(Expo))
        d_alpha = split(dat_, list(dat_$Year, dat_$Age_grp2, dat_$Expo)) %>%
            lapply(., function(x){dist(x[[aa]], method = "manhattan") %>% as.numeric}) %>% do.call("c", .) %>% data.frame(delta_alpha = .) %>% rownames_to_column(var = "Tmp") %>% separate(Tmp, c("Year", "Age_grp2", "Expo"), sep = "\\.") %>% mutate(Expo = gsub("\\d+", "", Expo), Age_grp2 = factor(Age_grp2, levels = c("Age_0-3", "Age_3-8", "Age_8-18", "Adults")))
        d_alpha.test = d_alpha %>% group_by(Expo, Age_grp2) %>% wilcox_test(delta_alpha ~ Year) %>% add_xy_position(x = "Age_grp2", dodge = 0.7) %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "fdr") %>% round(digits = 3), p.adj.signif = ifelse(p.adj<=0.1, p.adj, "ns"))
        
        ggplot(d_alpha, aes(x = Age_grp2, y = delta_alpha)) +
            geom_quasirandom(size = 1.5, aes(fill = Year), dodge.width = 0.7, shape = 21, alpha = 0.5) +
            geom_boxplot(width = 0.5, aes(color = Year), position = position_dodge(width = 0.7), fill = NA, size = 0.3, outlier.alpha = 0) +
            stat_pvalue_manual(d_alpha.test, label = "p.adj", tip.length = 0, size = 8/.pt) +
            facet_grid(.~Expo) +
            scale_fill_manual(values = c("#80b1d3", "#fb8072")) +
            scale_color_manual(values = c("black", "black")) +
            theme_jw(base_size = 8) + theme(aspect.ratio = 0.8) +
            labs(y = paste0("Absolute value \u0394(", aa, ")")) + ggtitle(bb)
        set_panel_size(file = paste0("fig_tmp/delta_alpha_year_", aa, "_", bb, ".pdf"), width = unit(3, "in"), height = unit(2.4, "in"))
    }
}
```


